\documentclass[11pt]{article}

% ── Packages ──────────────────────────────────────────────────────────
\usepackage[T1]{fontenc}
\usepackage[margin=2.5cm]{geometry}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{array,tabularx}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[final]{microtype}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,calc,decorations.pathreplacing}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage{enumitem}

% ── Theorem environments ──────────────────────────────────────────────
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{plain}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{lemma}[definition]{Lemma}
\newtheorem{proposition}[definition]{Proposition}
\newtheorem{conjecture}[definition]{Conjecture}
\theoremstyle{remark}
\newtheorem{remark}[definition]{Remark}
\newtheorem{invariant}[definition]{Invariant}

% ── Custom commands ───────────────────────────────────────────────────
\newcommand{\Ostr}{\mathcal{O}}          % Order strength
\newcommand{\Clvl}{\mathcal{C}}          % Closure level
\newcommand{\Cdep}{\mathcal{C}_{\mathrm{dep}}}
\newcommand{\Ccoh}{\mathcal{C}_{\mathrm{coh}}}
\newcommand{\Cfr}{\mathcal{C}_{\mathrm{fr}}}
\newcommand{\Cdepcert}{\mathcal{C}^{\mathrm{cert}}_{\mathrm{dep}}}
\newcommand{\Ccohcert}{\mathcal{C}^{\mathrm{cert}}_{\mathrm{coh}}}
\newcommand{\Ffin}{\mathcal{F}}          % Finality function
\newcommand{\Uord}{U_{\mathrm{order}}}   % Order utility
\newcommand{\Rcls}{R_{\mathrm{closure}}} % Closure risk
\newcommand{\Frag}{\mathrm{FragilityCoupling}}
\newcommand{\Phifn}{\Phi}               % Advisory mining functional
\newcommand{\Certcls}{\mathsf{Cert}_{\mathrm{cls}}}
\newcommand{\Dag}{\mathcal{G}}          % DAG graph
\newcommand{\Rset}{\mathbb{R}}          % Reals
\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{Y}{>{\raggedright\arraybackslash}X}
\newcommand{\codepath}[1]{\path{#1}}

\definecolor{blueblock}{RGB}{70,130,180}
\definecolor{redblock}{RGB}{205,92,92}
\definecolor{closurecolor}{RGB}{60,150,80}

% ── Typography and float tuning ───────────────────────────────────────
\linespread{1.03}
\allowdisplaybreaks[2]
\setlength{\emergencystretch}{2em}
\setlength{\textfloatsep}{14pt plus 2pt minus 4pt}
\setlength{\floatsep}{12pt plus 2pt minus 2pt}
\setlength{\intextsep}{12pt plus 2pt minus 2pt}
\captionsetup{font=small,labelfont=bf,skip=6pt}
\setlist[itemize,enumerate,description]{topsep=4pt,itemsep=2pt,parsep=0pt,partopsep=0pt}
\urlstyle{same}
\raggedbottom

% ── Metadata ──────────────────────────────────────────────────────────
\title{\textbf{PHASMDAG-Core: Closure-Certified Finality for High-Throughput DAG Consensus}\\[6pt]
{\large RobustDAG and the Separation of Ordering from Strong Finality\\
in Proof-of-Work BlockDAGs}}
\author{Arthur Zhang}
\date{April 2026}

\begin{document}
\maketitle

% =====================================================================
\begin{abstract}
This paper sets out a preliminary formulation of \emph{PHASMDAG-Core}, a
conservative high-throughput DAG protocol that may also be described, at the
level of design philosophy, as \emph{RobustDAG}.
The proposal is motivated by the contention that existing GhostDAG-family
designs conflate three distinct objects: \emph{ordering}, \emph{dependency
closure}, and \emph{finality}.
Recent work by Wu, Wei, Zhang, and Jiang\,(NDSS\,2024) suggests that the
\emph{Decoupling Assumption}---a prerequisite for the security analyses of the
major DAG-based PoW protocols---may fail under sufficiently high throughput,
owing to Block Jam and Late-Predecessor effects.
The present proposal therefore preserves a stable GhostDAG-style ordering core
with fixed consensus parameters, but adds a closure-aware finality layer in
which strong confirmation is granted only after deterministic,
work-weighted closure-certificate conditions hold over a bounded support
window.
PHASMDAG-Core treats closure as a monotonic finality certificate rather than a
volatile ordering weight.
Immature closure may delay strong confirmation, but it must not by itself
redirect canonical ordering.
This conservative asymmetry is intended to bound the payoff of closure
manipulation: adversarially delayed closure evidence may degrade latency, but
should not become a cheap consensus-steering lever.
Bi-objective closure-risk ordering is retained only as an experimental
extension for future work, not as the core rule of the proposed protocol.
The paper develops this v1 design sketch, states the principal theorem target
as \emph{certificate coherence} across honest nodes, and sets out a simulation
and implementation roadmap rather than a completed security proof.
\end{abstract}

\noindent{\small\textbf{Keywords:} DAG consensus; GhostDAG; closure-certified finality;
late predecessor; block jam; semantic separation; finality semantics}

% =====================================================================
\section{Introduction}\label{sec:intro}

DAG-based Proof-of-Work consensus protocols---SPECTRE~\cite{sompolinsky2016spectre},
PHANTOM/GhostDAG~\cite{sompolinsky2018phantom,sompolinsky2021ghostdag},
OHIE~\cite{yu2021ohie}, and Prism~\cite{bagaria2019prism}---represent the most
promising direction for scaling Nakamoto consensus beyond the throughput ceiling
imposed by chain-style protocols.
By allowing blocks to reference multiple parents and thereby capture concurrency,
these protocols achieve block rates that far exceed the network delay bound $D$.

However, Wu et al.~\cite{wu2024security} proved that this performance gain comes
with a previously unaccounted security cost.
Their \emph{Congestible Blockchain Model}\,(CBM) shows that when block production
rate $f$ is high relative to network capacity, two phenomena emerge:
\begin{enumerate}[nosep]
    \item \textbf{Block Jam}\,(BJ): when the total size of priority blocks in a
    time window exceeds the network's relay capacity, these blocks cannot
    propagate within the assumed delay~$D$.
    \item \textbf{Late Predecessor}\,(LP): a block's propagation completion does
    not imply its acceptability---a block can only be accepted after \emph{all}
    its predecessors have been received and validated.
    If a predecessor arrives later than its successor, the successor's actual
    delay ($\Delta_{\mathrm{actual}} = \Delta_{\mathrm{prop}} +
    \Delta_{\mathrm{proc}}$) exceeds its propagation delay alone.
\end{enumerate}

Together, BJ and LP violate the \emph{Decoupling
Assumption}~\cite{wu2024security}: the assumption that priority blocks can
propagate within fixed short delay $D$ \emph{and} be immediately accepted and
built upon.  All existing security proofs for DAG-based PoW protocols rely on
this assumption.

\paragraph{Contribution summary.}
\begin{enumerate}[nosep]
    \item \textbf{Root-cause diagnosis.}
    It is argued that the fundamental limitation of GhostDAG-family protocols
    lies not merely in the attackability of relay paths, but in the requirement
    that a single structural classifier should simultaneously represent
    canonical order, safe acceptability, and finality confidence.
    This conflation is largely invisible under optimistic network assumptions,
    but becomes a security liability under CBM.

    \item \textbf{Protocol design.}
    A candidate protocol design is set out in which closure is introduced as a
    first-class \emph{finality-certification} state, while the canonical
    ordering core is deliberately left structurally stable.
    PHASMDAG-Core preserves a GhostDAG-style blue-work lineage for tentative
    ordering, but grants strong confirmation only after a monotonic,
    work-weighted closure certificate has been issued.

    \item \textbf{Preliminary security analysis.}
    A conservative security framing is proposed in which LP/BJ stress is
    intended to delay strong finality before it can silently weaken safety
    semantics.
    The central claim is not that closure manipulation is eliminated, but that
    its payoff is bounded by refusing to let immature closure directly steer
    canonical order.

    \item \textbf{Formal modelling and verification roadmap.}
    A compositional formal model, a simulation-based evaluation methodology,
    and a phased verification programme are delineated around the theorem
    target of certificate coherence.
\end{enumerate}

\paragraph{On the designation ``PHASM''.}
The designation \emph{PHASMDAG-Core} is intended to evoke
\emph{phase-separated semantics}: order, closure, and finality are treated as
distinct protocol phases rather than being collapsed into a single structural
score.

\paragraph{Claim status.}
This document is intentionally written as a protocol proposal rather than as a
completed proof artifact.
\textbf{Definitions} specify protocol intent.
\textbf{Lemmas} and \textbf{Propositions} are preliminary and hold only under
the modelling assumptions stated around them.
\textbf{Conjectures} identify unresolved validation targets.
The most important unresolved target is \emph{certificate coherence}: whether raw
closure observations and certificate-side thresholds converge sufficiently
across honest nodes to support consistent strong confirmation.

\paragraph{Design principles.}
\begin{enumerate}[nosep]
    \item \textbf{Immature closure delays strong confirmation; it does not directly
    penalise canonical order.}
    \item \textbf{Closure is a monotonic certification layer, not a volatile
    ordering weight.}
    \item \textbf{The v1 protocol is conservative}: fixed $K$, fixed support
    window $H$, no dynamic $\lambda$, no dynamic $K$, no closure bounty, and no
    telemetry-driven local consensus parameters.
\end{enumerate}
The guiding intuition is simple: \emph{a robust high-throughput DAG should
parallelise ordering, not finality.  Strong finality is serialized through
dependency closure.}

% =====================================================================
\section{Related Work and Theoretical Context}\label{sec:background}

\subsection{Throughput, Security, and Delay in PoW Consensus}

The tension between throughput and network delay predates DAG-based consensus.
Sompolinsky and Zohar's high-rate Bitcoin analysis~\cite{sompolinsky2013secure}
already emphasised that short inter-block intervals and bounded delay cannot be
treated independently.
Formal longest-chain analyses subsequently made this dependence explicit
through common-prefix, chain-growth, and chain-quality style guarantees under
bounded-delay assumptions~\cite{garay2015backbone}.
Zhang and Preneel~\cite{zhang2019common} later unified these perspectives into
a common-metrics framework, making it possible to compare protocol families
under aligned security notions rather than under protocol-specific metrics.

Two considerations may be taken from that literature.
First, performance claims in PoW systems are meaningful only relative to an
explicit delay model.
Second, security arguments become fragile when propagation, acceptance, and
mining eligibility are silently conflated.

\subsection{DAG-Based PoW Ordering Protocols}

The DAG-based PoW literature can be organised by the manner in which
concurrency is converted into order:
\begin{description}[nosep]
    \item[SPECTRE~\cite{sompolinsky2016spectre}] provides fast confirmation
    through pairwise voting, but does not aim to maintain a global total order.

    \item[PHANTOM~\cite{sompolinsky2018phantom}] formulates ordering as a
    $k$-cluster identification problem, thereby making explicit the relation
    between concurrency and anticone structure.

    \item[GhostDAG~\cite{sompolinsky2021ghostdag}] replaces exact clustering
    with a greedy $K$-cluster approximation, yielding a practical structural
    order rule based on blue-set growth and blue-work accumulation.

    \item[OHIE~\cite{yu2021ohie}] separates transaction handling from
    structural ordering through a numbered-chain construction.

    \item[Prism~\cite{bagaria2019prism}] decomposes consensus into proposal,
    voting, and committee roles, thereby achieving high throughput through
    role separation rather than through a single ordering object.
\end{description}

The present paper differs from these systems in one principal respect.
The distinctive feature is not the use of a DAG per se, but the explicit
separation of \emph{order}, \emph{closure maturity}, and \emph{finality} as
distinct protocol-level semantic objects.

\subsection{Congestion, Delayed Acceptability, and Strategic Deviation}

Wu et al.~\cite{wu2024security} identify the strongest current challenge to the
security analysis of high-rate DAG protocols.
Their Congestible Blockchain Model shows that propagation completion need not
coincide with acceptability once block jam and late-predecessor effects are
present.
This result is directly relevant here because it shifts the analytical focus
from pure relay delay to dependency visibility and processing delay.

The incentive literature is also relevant.
Strategic withholding attacks were first articulated clearly in the selfish
mining literature~\cite{eyal2013majority}, where delayed publication changes
the effective security margin even without violating PoW validity.
Closure manipulation in PHASMDAG-Core is not identical to selfish mining, but it
belongs to the same broad family of attacks in which timing asymmetry and
selective release alter the consensus-visible state.
Any academically complete treatment of a closure-aware protocol must therefore
address strategic publication incentives, not merely structural safety.

\subsection{The Congestible Blockchain Model}

Wu et al.~\cite{wu2024security} introduce the Congestible Blockchain Model
(CBM), which relaxes the Uniform Delay Bounded Model (UDBM) used in prior
security proofs.

\begin{definition}[CBM, informal]\label{def:cbm}
Under CBM, each block $B$ experiences a block-specific propagation delay upper
bound $D_B$ that may exceed the nominal network delay $D$.  Specifically,
$D_B$ depends on:
\begin{enumerate}[nosep]
    \item The bandwidth utilisation at the time of $B$'s propagation
    (Block Jam).
    \item The arrival status of $B$'s predecessors at each receiving node
    (Late Predecessor).
\end{enumerate}
The actual delay for block $B$ at node $v$ is:
\begin{equation}\label{eq:actual_delay}
\Delta_{\mathrm{actual}}(B, v)
= \Delta_{\mathrm{prop}}(B, v) + \Delta_{\mathrm{proc}}(B, v)
\end{equation}
where $\Delta_{\mathrm{proc}}(B, v)$ is the processing delay caused by late
predecessors of $B$ that have not yet arrived at~$v$.
\end{definition}

\begin{theorem}[Security degradation under CBM~{\cite{wu2024security}}]\label{thm:cbm}
For a DAG-based PoW protocol under LP attack with $s$ node groups, the chain
growth rate of affected blocks is:
\begin{equation}\label{eq:chain_growth_cbm}
g = (1 - \sigma)\,\gamma\,f, \qquad
\gamma = \frac{\alpha}{1 + \alpha f\,\mathbb{E}[\Delta]}
\end{equation}
where $\sigma$ is the attacker's hash rate fraction, $\alpha = 1 - \sigma$ is
the honest fraction, and $\mathbb{E}[\Delta]$ is the expected actual delay.
The security threshold $\beta < \gamma$ shrinks as $\mathbb{E}[\Delta]$
increases.
\end{theorem}

\subsection{The Decoupling Assumption}

\begin{definition}[Decoupling Assumption~{\cite{wu2024security}}]\label{def:decoupling}
A DAG-based PoW protocol satisfies the Decoupling Assumption if there exists a
class of priority blocks such that:
\begin{enumerate}[nosep]
    \item Priority blocks propagate to all honest nodes within fixed delay~$D$.
    \item Upon receipt, honest nodes can immediately accept priority blocks and
    build upon them.
\end{enumerate}
\end{definition}

\noindent Block Jam negates condition~(1); Late Predecessor negates
condition~(2).  Zhang and Preneel~\cite{zhang2019common} further show that this
assumption is not merely a technical convenience---it is \emph{necessary} for
the existing proof structure of GhostDAG-family protocols, because the
$K$-parameter derivation fundamentally requires that the anticone of any block
is bounded by $K$ with high probability, which in turn requires that blocks are
accepted within $D$.

% =====================================================================
\section{Problem Statement: The Semantic Conflation}\label{sec:problem}

\subsection{Three Conflated Semantics}

The following conflation appears in GhostDAG-family protocols:

\begin{definition}[Semantic Conflation]\label{def:conflation}
GhostDAG-family protocols use a single structural object---the blue score /
$K$-cluster colouring---to simultaneously represent three distinct semantic
roles:
\begin{enumerate}[nosep]
    \item \textbf{Order}: the canonical position of a block in the DAG's
    preferred history.
    \item \textbf{Acceptability}: whether a block's dependency world is
    sufficiently complete for honest miners to safely build upon it.
    \item \textbf{Finality}: the economic confidence that a block will not be
    reorged.
\end{enumerate}
Under UDBM with low $f$, these three roles approximately coincide.
Under CBM with high $f$, they diverge: a block can be structurally well-ordered
(high blue score) yet have an immature dependency world (low acceptability), and
acceptability does not guarantee finality.
\end{definition}

This conflation is not merely an implementation oversight; it is a
\emph{mathematical} property of protocols whose consensus functional is purely
structural:
\begin{equation}\label{eq:ghostdag_objective}
\text{GhostDAG:} \quad \max_{S}\; \Uord(S)
\qquad\text{(structural concurrency only)}
\end{equation}

When $\Uord(S)$ is the sole objective, there is no mechanism within the
protocol to distinguish between ``structurally legal but closure-fragile''
support and ``structurally legal and closure-mature'' support.
The weakness of GhostDAG under CBM is therefore not merely that the inherited
$K$ parameter may become poorly calibrated.
The deeper difficulty is that structural ordering is tacitly treated as a
proxy for dependency acceptability and finality.
PHASMDAG-Core does not attempt to repair this by making ordering ever more
adaptive.
It instead refuses to let structural order stand in for strong finality.

\subsection{Why Non-Consensus Mitigations Are Necessary But Insufficient}

Several implementation-level mitigations are natural responses to the LP/BJ
challenge:
\begin{itemize}[nosep]
    \item \textbf{Header-First Propagation}: decouple header acceptance from body
    completion, allowing GhostDAG to progress on header closure alone.
    \item \textbf{Dependency-Aware Relay Priority}: prioritise relay of blocks
    that unblock the most pending successors.
    \item \textbf{Telemetry}: expose dependency-completion and orphan-age
    metrics.
    \item \textbf{Adaptive Safety Margins}: locally increase conservatism when
    network stress is detected.
\end{itemize}

These mitigations are valuable and should be pursued whenever possible.
However, they share a common limitation: they protect a protocol whose
core mathematical object remains structurally blind to closure.

\begin{proposition}\label{prop:engineering_ceiling}
Engineering mitigations that preserve the purely structural consensus objective
$\max\,\Uord(S)$ can reduce the magnitude of closure fragility but cannot
detect it, certify it, or let it gate strong finality within the consensus
protocol itself.
Therefore, the security guarantee remains dependent on the Decoupling
Assumption, albeit at a lower effective delay.
\end{proposition}

Header-first, for example, narrows the LP attack surface from the body layer to
the header layer, but does not eliminate it.  The question ``is this block's
dependency world mature enough for strong finality?'' remains unanswerable
within the protocol---it is answered only by heuristics in the implementation
layer.
This semantic mismatch is summarised visually in \Cref{fig:semantic_gap}.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    stage/.style={draw, rounded corners=3pt, minimum width=2.1cm,
                  minimum height=0.82cm, align=center, font=\footnotesize,
                  line width=0.7pt, fill=white},
    ord/.style={stage, draw=blueblock!75!black, fill=blueblock!8},
    cls/.style={stage, draw=closurecolor!70!black, fill=closurecolor!8},
    fin/.style={stage, draw=redblock!75!black, fill=redblock!8},
    arr/.style={-{Stealth[length=2.2mm]}, line width=0.9pt, draw=black!70},
    warn/.style={-{Stealth[length=2.2mm]}, line width=0.9pt,
                 draw=redblock!75!black, dashed},
    note/.style={font=\scriptsize, align=center, text=black!72,
                 fill=white, inner sep=1.2pt},
    group/.style={font=\scriptsize\bfseries, text=black!65}
]
\node[group, anchor=east] at (-0.25,1.0) {Conflated};
\node[ord] (gobs) at (1.1,1.0) {Observed\\DAG};
\node[ord] (gord) at (4.0,1.0) {Order};
\node[fin] (gfin) at (6.9,1.0) {Finality};
\draw[arr] (gobs.east) -- (gord.west);
\draw[warn] (gord.east) -- (gfin.west);
\node[note, text=redblock!75!black] at (5.45,1.67) {implicit closure\\judgement};
\node[note, text=redblock!75!black] at (5.45,0.28) {LP gap};

\node[group, anchor=east] at (-0.25,-1.0) {Separated};
\node[ord] (pobs) at (1.1,-1.0) {Observed\\DAG};
\node[ord] (pord) at (3.55,-1.0) {Order};
\node[cls] (pcls) at (6.0,-1.0) {Closure};
\node[fin] (pfin) at (8.45,-1.0) {Finality};
\draw[arr] (pobs.east) -- (pord.west);
\draw[arr] (pord.east) -- (pcls.west);
\draw[arr] (pcls.east) -- (pfin.west);
\node[note] at (6.0,-1.78) {closure becomes consensus-visible\\before strong confirmation};
\end{tikzpicture}
\caption{Why an explicit closure layer is needed.  Under LP attack, structural
ordering can progress before the dependency world is mature enough for strong
confirmation.  PHASMDAG-Core inserts closure as an explicit protocol stage rather
than leaving that judgement to implementation heuristics.}
\label{fig:semantic_gap}
\end{figure}

\subsection{Formal Problem Definition}

\begin{definition}[The PHASMDAG-Core Problem]\label{def:problem}
Design a DAG consensus protocol that:
\begin{enumerate}[nosep]
    \item preserves GhostDAG-style high-throughput structural ordering,
    \item exposes dependency closure as a protocol-visible finality condition,
    \item prevents closure evidence from becoming a cheap
    ordering-manipulation vector,
    \item grants strong confirmation only after closure certification, and
    \item degrades under LP/BJ by delaying strong finality rather than
    silently weakening safety semantics.
\end{enumerate}
\end{definition}

% =====================================================================
\section{Protocol Design}\label{sec:design}

\subsection{Three Semantic Layers}

PHASMDAG-Core reasons over three explicit semantic layers:

\begin{description}[nosep]
    \item[Layer A --- Order Layer] determines structural support and preferred
    ordering relations in the DAG.  In PHASMDAG-Core-$v_1$, this layer is fixed to a
    GhostDAG-style \emph{blue-work lineage}; alternative order cores are future
    variants, not part of the base protocol.

    \item[Layer B --- Closure Layer] measures whether a block and its dependency
    world are sufficiently mature, stable, and coherently visible across honest
    nodes, and promotes only sufficiently stable evidence to a monotonic
    certificate.  This is the additional layer absent from GhostDAG-family
    protocols.

    \item[Layer C --- Finality Layer] derives confirmation/finality from both
    order and certified closure, rather than from order alone.
\end{description}

The relationship between layers is illustrated in \Cref{fig:layers}.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    layer/.style={draw, rounded corners=3pt, minimum width=8.8cm,
                  minimum height=0.95cm, font=\footnotesize, align=left,
                  line width=0.7pt, inner xsep=7pt, fill=white},
    ord/.style={layer, draw=blueblock!75!black, fill=blueblock!8},
    cls/.style={layer, draw=closurecolor!70!black, fill=closurecolor!8},
    fin/.style={layer, draw=redblock!75!black, fill=redblock!8},
    arr/.style={-{Stealth[length=2.2mm]}, line width=0.9pt, draw=black!70}
]
\node[ord] (A) at (0,0)
    {\textbf{Layer A: Order}\quad $\Ostr(B)$\\blue work, preferred chain, canonical rank};
\node[cls] (B) at (0,-1.8)
    {\textbf{Layer B: Closure}\quad $\Clvl(B),\,\Certcls(B)$\\raw evidence, discounting, certificate};
\node[fin] (C) at (0,-3.6)
    {\textbf{Layer C: Finality}\quad $\Ffin(B)$\\grade unlocked by order, certificate, and depth};
\draw[arr] (A) -- (B);
\draw[arr] (B) -- (C);
\draw[arr, dashed, bend left=28] (A.east) to (C.east);
\end{tikzpicture}
\caption{The three semantic layers of PHASMDAG-Core.  The Order layer feeds the
Closure layer; both jointly determine the Finality layer.  The dashed arrow
indicates that Order also feeds Finality directly, but strong confirmation is
not unlocked until the Closure layer issues a certificate.}
\label{fig:layers}
\end{figure}

\subsection{Core Quantities}

For each block $B$ in the observed DAG $\Dag = (V, E)$,
PHASMDAG-Core tracks four objects:
the structural order score $\Ostr(B)$,
the raw closure observation $\Clvl(B)$,
the monotonic closure certificate $\Certcls(B)$,
and the resulting finality grade $\Ffin(B)$.

\begin{definition}[Order Strength]\label{def:order_strength}
For PHASMDAG-Core-$v_1$, the \emph{order strength} of block $B$ is the block's
position on the GhostDAG blue-work lineage:
\begin{equation}\label{eq:order_strength}
\Ostr(B) := \mathrm{BlueWork}(B), \qquad \Ostr(B) \in \Rset_{\ge 0}
\end{equation}
a structural support quantity measuring how strongly the DAG structure supports
placing $B$ in the canonical history.
The canonical order rank $r(B)$ is the deterministic rank induced by this
blue-work order core.
\end{definition}

\begin{definition}[Relevant Support Window]\label{def:support_window}
Let $H \in \mathbb{N}$ be a protocol parameter.
The \emph{relevant support window} of block $B$ is
\begin{equation}\label{eq:support_window}
\mathcal{W}_H(B) :=
\{D \in \mathrm{desc}(B) : 0 < r(D) - r(B) \le H\}.
\end{equation}
All closure metrics in PHASMDAG-Core-$v_1$ are computed over $\mathcal{W}_H(B)$ so
that they are bounded, deterministic, and auditable.
\end{definition}

\begin{definition}[Closure Evidence Sub-Metrics]\label{def:closure_submetrics}
Let $W(B)$ denote the PoW work attributed to block $B$.
For each $D \in \mathcal{W}_H(B)$, let $\mathbf{1}_{\mathrm{closed}}(D)=1$ if
all protocol-required predecessor objects of $D$ have been observed and
validated at the evaluating node, and $0$ otherwise.
Let $c^\star(B)$ denote the canonical child of $B$ on the blue-work lineage,
and let $\pi_B(D)$ denote the first child of $B$ on the path from $B$ to $D$.
Let $\mathrm{Front}_H(B)$ be the set of maximal blocks in $\mathcal{W}_H(B)$,
and let $f^\star(B) \in \mathrm{Front}_H(B)$ be the frontier tip that extends
$c^\star(B)$ when such a tip exists.
The sub-metrics are defined by:
\begin{align}
\Cdep(B)
&=
\frac{\sum_{D \in \mathcal{W}_H(B)} W(D)\,\mathbf{1}_{\mathrm{closed}}(D)}
{\sum_{D \in \mathcal{W}_H(B)} W(D)}
\label{eq:cdep} \\
\Ccoh(B)
&=
\frac{\sum_{D \in \mathcal{W}_H(B)} W(D)\,\mathbf{1}[\pi_B(D)=c^\star(B)]}
{\sum_{D \in \mathcal{W}_H(B)} W(D)}
\label{eq:ccoh} \\
\Cfr(B)
&=
\frac{W(f^\star(B))}
{\sum_{F \in \mathrm{Front}_H(B)} W(F)}
\label{eq:cfr}
\end{align}
with the convention that each ratio is $0$ when its denominator is $0$.
Define also the support-window work evidence
\begin{equation}\label{eq:work_window}
\mathrm{Work}_H(B) := \sum_{D \in \mathcal{W}_H(B)} W(D).
\end{equation}
\end{definition}

\begin{definition}[Raw Closure Observation]\label{def:closure_level}
The \emph{raw closure observation} of block $B$ is the deterministic
aggregation of the three closure sub-metrics:
\begin{equation}\label{eq:closure_level}
\Clvl(B) = g_{v_1}\!\bigl(\Cdep(B),\,\Ccoh(B),\,\Cfr(B)\bigr)
\end{equation}
where PHASMDAG-Core-$v_1$ uses the weighted geometric mean
\begin{equation}\label{eq:closure_level_v1}
g_{v_1}(x,y,z) = x^{\eta_{\mathrm{dep}}}
y^{\eta_{\mathrm{coh}}}
z^{\eta_{\mathrm{fr}}},
\qquad
\eta_{\mathrm{dep}}+\eta_{\mathrm{coh}}+\eta_{\mathrm{fr}}=1.
\end{equation}
This aggregation makes a low value in any one dimension visibly reduce the
overall raw closure score.
For typographic continuity, the paper writes $\Clvl(B)$ for this node-local
quantity.
It is intentionally treated as an intermediate observation rather than as a
direct trigger for strong confirmation.
\end{definition}

\begin{definition}[Closure Certificate]\label{def:closure_certificate}
For PHASMDAG-Core-$v_1$, the certificate-side closure components are
\begin{align}
\Cdepcert(B)
&= 
\frac{\sum_{D \in \mathcal{W}_H(B)} W(D)\,\mathbf{1}_{\mathrm{closed}}(D)}
{\mathrm{Work}_H(B)}
\label{eq:cdep_cert} \\
\Ccohcert(B)
&= 
\frac{\sum_{D \in \mathcal{W}_H(B)} W(D)\,\mathbf{1}[\pi_B(D)=c^\star(B)]}
{\mathrm{Work}_H(B)}.
\label{eq:ccoh_cert}
\end{align}
Define the certificate predicate
\begin{equation}\label{eq:valid_cert}
\mathrm{ValidCert}_{\mathcal{C}}(B)
:=
\bigl[\Cdepcert(B) \ge \theta_{\mathrm{dep}}\bigr]
\wedge
\bigl[\Ccohcert(B) \ge \theta_{\mathrm{coh}}\bigr]
\wedge
\bigl[\mathrm{Work}_H(B) \ge \theta_{\mathrm{work}}\bigr].
\end{equation}
Block $B$ is said to be \emph{closure-certified}, written $\Certcls(B)=1$,
iff $\mathrm{ValidCert}_{\mathcal{C}}(B)$ holds; once issued, the certificate
is monotonic.
\end{definition}

\begin{remark}\label{rem:closure_vs_delay}
$\Clvl(B)$ is \emph{not} a function of observed propagation delay.
It measures structural properties of the dependency world (completeness,
stability, coherence), not temporal observations.
This is a deliberate design choice: observed delays are node-specific and
vulnerable to topology bias~\cite{wu2024security}, whereas structural
completeness can, in principle, be determined deterministically from the DAG
structure that honest nodes eventually agree on.
Strong confirmation is therefore tied not to the live score $\Clvl(B)$ itself,
but to the monotonic event $\Certcls(B)=1$.
\end{remark}

\begin{remark}[Why $\Cfr(B)$ is demoted in $v_1$]\label{rem:closure_decomp}
\Cref{eq:cdep} and \Cref{eq:ccoh} are retained as the mandatory Grade~2
certificate inputs because they speak directly to dependency closure and
coherent descendant support.
By contrast, $\Cfr(B)$ is more exposed to frontier steering and local tip
instability.
PHASMDAG-Core therefore keeps $\Cfr(B)$ as an advisory or deep-settlement
quantity in $v_1$, rather than making it a brittle Grade~2 gate.
\end{remark}
\Cref{fig:closure_components} shows how these three quantities are read off from
the local descendant structure around a block.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    x=1cm, y=1cm,
    blk/.style={circle, minimum size=7.4mm, inner sep=0pt, draw=black!55,
                fill=white, line width=0.7pt, font=\scriptsize},
    focus/.style={blk, draw=blueblock!75!black, fill=blueblock!8},
    closed/.style={blk, draw=closurecolor!70!black, fill=closurecolor!8},
    unresolved/.style={blk, draw=redblock!75!black, fill=white, dashed},
    edge/.style={-{Stealth[length=2.2mm]}, line width=0.8pt, draw=black!55},
    pref/.style={-{Stealth[length=2.2mm]}, line width=0.9pt, draw=blueblock!75!black},
    boxlabel/.style={font=\scriptsize, align=left, text=black!78},
    key/.style={font=\scriptsize, text=black!72}
]
\draw[rounded corners=3pt, dashed, draw=black!45, line width=0.7pt]
    (1.1,-1.7) rectangle (6.75,1.7);
\fill[black!4, rounded corners=2pt] (5.45,-1.52) rectangle (6.6,1.52);
\node[key, anchor=south west] at (1.2,1.72) {support window $\mathcal{W}_H(B)$};
\node[key, anchor=south] at (6.02,1.55) {frontier};

\node[focus] (B) at (0,0) {$B$};
\node[focus] (c) at (2,0.9) {$c^\star$};
\node[blk]   (x) at (2,-0.9) {$x$};
\node[closed]     (d1) at (4,1.35) {$D_1$};
\node[unresolved] (d2) at (4,0.0) {$D_2$};
\node[closed]     (d3) at (4,-1.35) {$D_3$};
\node[closed]     (f1) at (6.05,1.35) {$f^\star$};
\node[unresolved] (f2) at (6.05,0.0) {$F_2$};
\node[closed]     (f3) at (6.05,-1.35) {$F_3$};

\draw[pref] (B) -- (c);
\draw[edge] (B) -- (x);
\draw[pref] (c) -- (d1);
\draw[pref] (c) -- (d2);
\draw[edge] (x) -- (d3);
\draw[pref] (d1) -- (f1);
\draw[pref] (d2) -- (f2);
\draw[edge] (d3) -- (f3);

\node[boxlabel, text width=4.25cm, anchor=west] at (7.35,1.0)
    {$\Cdep(B)$: fraction of support-window work that is already closed};
\node[boxlabel, text width=4.25cm, anchor=west] at (7.35,0.0)
    {$\Ccoh(B)$: work routed through the canonical child $c^\star(B)$};
\node[boxlabel, text width=4.25cm, anchor=west] at (7.35,-1.0)
    {$\Cfr(B)$: frontier mass concentrated on the preferred tip $f^\star(B)$};

\node[closed, minimum size=4.8mm] at (1.35,-2.18) {};
\node[key, anchor=west] at (1.65,-2.18) {closure-complete descendant};
\node[unresolved, minimum size=4.8mm] at (4.55,-2.18) {};
\node[key, anchor=west] at (4.85,-2.18) {unresolved support};
\end{tikzpicture}
\caption{Closure decomposition around block $B$.  The support window
$\mathcal{W}_H(B)$ supplies the auditable evidence set; $\Cdep(B)$ measures how
much of that work is closure-complete, $\Ccoh(B)$ measures how much support
flows through the canonical child $c^\star(B)$, and $\Cfr(B)$ measures whether
frontier work is concentrated on the preferred tip $f^\star(B)$.
In PHASMDAG-Core-$v_1$, the first two quantities are mandatory Grade~2 inputs,
while $\Cfr(B)$ is reserved for diagnostics or deep settlement.}
\label{fig:closure_components}
\end{figure}

\subsection{Closure-Certified Finality, Not Closure-Weighted Ordering}

\begin{definition}[Order-Preservation Principle]\label{def:order_preserving}
Canonical order in PHASMDAG-Core-$v_1$ is determined by the GhostDAG-style
blue-work order core alone.
Immature closure may delay strong confirmation, but it must not by itself
redirect canonical ordering.
\end{definition}

\begin{definition}[Certificate-Gated Finality]\label{def:finality}
Let $D(B)$ denote the settlement-depth or descendant-work condition used for
deep settlement.
The protocol exposes three grades:
\begin{align}
\mathsf{Grade}_1(B)
&\iff
\Ostr(B) \ge \tau_{\mathrm{ord}},
\label{eq:grade1} \\
\mathsf{Grade}_2(B)
&\iff
\mathsf{Grade}_1(B) \wedge \mathrm{ValidCert}_{\mathcal{C}}(B),
\label{eq:grade2} \\
\mathsf{Grade}_3(B)
&\iff
\mathsf{Grade}_2(B) \wedge D(B) \ge \tau_{\mathrm{settle}}
\wedge \bigl[\Cfr(B) \ge \theta^{\mathrm{deep}}_{\mathrm{fr}}\bigr]_{\mathrm{optional}}.
\label{eq:grade3}
\end{align}
The finality grade $\Ffin(B)$ is the greatest $i \in \{0,1,2,3\}$ such that
$\mathsf{Grade}_i(B)$ holds, with $\mathsf{Grade}_0$ denoting the absence of
protocol-level confirmation.
\end{definition}

PHASMDAG-Core treats closure as a monotonic finality certificate rather than a
volatile ordering weight.
Strong finality is serialized through dependency closure.
This is the protocol's central asymmetry:
positive closure evidence unlocks stronger confirmation, whereas missing
closure evidence does not directly punish canonical order.

\begin{description}[nosep]
    \item[Grade 1 --- Tentative Ordering]
    The block is structurally well placed in the GhostDAG-style order core.

    \item[Grade 2 --- Closure-Certified Strong Confirmation]
    The block satisfies the order threshold and has a valid closure
    certificate.
    Raw closure observations alone are insufficient.

    \item[Grade 3 --- Deep Settlement]
    The block is already Grade~2 and additionally satisfies a deep-settlement
    condition.
    Frontier stability may enter here as an optional strengthening, but it is
    not required for Grade~2 in $v_1$.
\end{description}

This replaces the traditional ``one confirmation number fits all'' model with
semantically differentiated confirmation objects.
The geometry of these grades is illustrated in \Cref{fig:grades_plane}.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    x=1cm, y=1cm,
    regionlabel/.style={font=\footnotesize, align=center},
    axislabel/.style={font=\scriptsize, text=black!75},
    callout/.style={draw, rounded corners=3pt, fill=redblock!8,
                    line width=0.7pt, align=center, font=\footnotesize}
]
\fill[black!4] (0,0) rectangle (2.35,4.0);
\fill[blueblock!8] (2.35,0) rectangle (6.05,2.0);
\fill[closurecolor!9] (2.35,2.0) rectangle (6.05,4.0);

\draw[->, line width=0.9pt] (0,0) -- (6.45,0)
    node[below right, axislabel] {order strength $\Ostr(B)$};
\draw[->, line width=0.9pt] (0,0) -- (0,4.3)
    node[left, axislabel] {certificate validity / settlement condition};
\draw[dashed] (2.35,0) -- (2.35,4.0) node[below, axislabel] {$\tau_{\mathrm{ord}}$};
\draw[dashed] (0,2.0) -- (6.05,2.0) node[left, axislabel] {$\mathrm{ValidCert}_{\mathcal{C}}$};

\node[regionlabel] at (1.15,2.05) {No confirmation\\or weak ordering};
\node[regionlabel] at (4.2,0.95) {Grade 1\\Tentative ordering};
\node[regionlabel] at (4.15,3.15) {Grade 2\\Strong confirmation};

\draw[-{Stealth[length=2.2mm]}, line width=0.9pt, draw=redblock!80!black]
    (6.05,3.0) -- (6.45,3.0);
\node[callout, minimum width=2.85cm, minimum height=1.1cm] at (8.1,3.0)
    {Grade 3\\Deep settlement\\requires sufficient $D(B)$};
\end{tikzpicture}
\caption{Schematic confirmation-grade geometry.
A robust high-throughput DAG should parallelise ordering, not finality.
Grade~2 is unlocked by certificate validity rather than by a volatile local
closure score; Grade~3 adds deep settlement on top of Grade~2.}
\label{fig:grades_plane}
\end{figure}

\subsection{Diagnostic Closure-Risk Metrics and Optional Extensions}

\begin{definition}[Diagnostic Closure-Risk Functional]\label{def:rclosure}
For a candidate set $S$, define the diagnostic closure-fragility metric
\begin{equation}\label{eq:rclosure}
\Rcls(S) =
a\sum_{B\in S}(1-\Cdep(B))
{}+ b\sum_{B\in S}(1-\Ccoh(B))
{}+ c\sum_{B\in S}(1-\Cfr(B))
{}+ d\cdot \Frag(S)
\end{equation}
where $a,b,c,d \ge 0$ and
\begin{equation}\label{eq:unresolved_set}
\mathcal{U}(B) :=
\{D \in \mathcal{W}_H(B) : \mathbf{1}_{\mathrm{closed}}(D)=0\},
\end{equation}
\begin{equation}\label{eq:fragility_coupling}
\Frag(S)
:=
\frac{1}{|P_2(S)|}
\sum_{\{B_i,B_j\}\in P_2(S)}
\frac{\sum_{D \in \mathcal{U}(B_i)\cap \mathcal{U}(B_j)} W(D)}
{\sum_{D \in \mathcal{U}(B_i)\cup \mathcal{U}(B_j)} W(D)},
\end{equation}
with value $0$ when the denominator is $0$.
\end{definition}

\begin{remark}[Status of $\Rcls$ and $\lambda$ in $v_1$]\label{rem:rclosure_audit}
In PHASMDAG-Core, $\Rcls$ is a diagnostic and simulation object in $v_1$, not
a hard canonical-ordering penalty.
It may still be useful for empirical comparison, reporting closure fragility,
or studying optional miner-local template preferences in simulations.
However, if $\Rcls$ directly affects canonical ordering, closure manipulation
becomes a consensus-steering problem.
PHASMDAG-Core therefore deliberately avoids hard $\Rcls$ penalties in the
deployed ordering rule.
\end{remark}

\begin{remark}[$\lambda$ is not consensus-critical in $v_1$]\label{rem:lambda}
PHASMDAG-Core has no dynamic $\lambda$, no telemetry-driven local $\lambda$,
and no per-node adaptive $\lambda$ in the consensus rule.
If a future extension studies a $\lambda$-bearing objective, it should be
understood as a separate protocol family that requires its own
incentive-compatibility analysis.
\end{remark}

\subsection{Confirmation Invariants}

\begin{invariant}[Tentative Ordering Bounded Churn]\label{inv:tentative}
There exists a protocol parameter $\rho_{\mathrm{tent}}$ such that once a block
$B$ attains Grade~1, any later honest view that still places $B$ in the
canonical order may change its rank by at most $\rho_{\mathrm{tent}}$.
Tentative ordering may move, but only within a bounded local reorder depth.
\end{invariant}

\begin{invariant}[Strong Confirmation Agreement]\label{inv:strong}
No two honest nodes may assign Grade~2 to conflicting blocks at the same
virtual position in the canonical order.
Strong confirmation is therefore a protocol object, not merely a UI label.
\end{invariant}

\begin{invariant}[Deep Settlement Tail]\label{inv:settlement}
There exists a protocol target $\varepsilon_{\mathrm{settle}}$ such that any
block assigned Grade~3 has rollback probability at most
$\varepsilon_{\mathrm{settle}}$ under the stated network and adversarial
assumptions.
\end{invariant}

% =====================================================================
\section{Security Analysis under CBM}\label{sec:security}

\subsection{Threat Model}

The CBM threat model of~\cite{wu2024security} is adopted with three layers:

\begin{enumerate}[nosep]
    \item \textbf{Analytical layer}: an abstract CBM attacker who can partition
    honest nodes into $s$ groups and selectively delay block delivery.
    \item \textbf{Realisable layer}: attacks implementable via BGP hijacking,
    Eclipse attacks, or strategic peer selection.
    \item \textbf{Implementation layer}: representative node behaviour under
    stress, including orphan-pool dynamics and header-body skew.
\end{enumerate}

\subsection{Closure Manipulation and Payoff Bounding}\label{sec:incentive}

This section is intentionally conservative.
It does \emph{not} claim that PHASMDAG-Core eliminates closure manipulation.
Rather, the design objective is to limit the payoff of such manipulation by
refusing to let immature closure directly steer canonical order.

\begin{proposition}[Payoff-bounding design intention]\label{prop:graceful}
Under the PHASMDAG-Core-$v_1$ design, an attacker may delay closure evidence
and thereby postpone Grade~2 confirmation.
However, because closure does not directly alter canonical order in $v_1$,
such manipulation is primarily a latency-griefing vector rather than an
ordering-redirection vector.
\end{proposition}

\begin{proof}[Proof sketch]
By \Cref{def:order_preserving}, canonical order depends only on the
GhostDAG-style order core and its fixed consensus parameters.
Closure evidence enters at the Grade~2 boundary through
\Cref{eq:valid_cert,eq:grade2}, not through the ordering rule itself.
Withholding or selectively releasing predecessor information may therefore
delay the moment at which $\mathrm{ValidCert}_{\mathcal{C}}(B)$ becomes true,
but it does not by itself re-rank a block that is already canonically ordered.
\end{proof}

The protocol deliberately bounds the payoff of closure manipulation:
adversarially delayed closure evidence can degrade latency, but should not
become a cheap consensus-steering lever.

\paragraph{Predecessor withholding.}
An attacker may withhold a predecessor block $B^\star$ so as to delay
dependency closure for descendants that rely upon it.
In GhostDAG, such delay risks being mistaken for weak structural support.
In PHASMDAG-Core, the intended effect is narrower:
the victim's block may remain in Grade~1 for longer, but tentative canonical
order is not thereby revoked.

\paragraph{Timed closure release.}
An attacker may release previously withheld closure evidence at a moment chosen
to accelerate a preferred branch's transition from Grade~1 to Grade~2.
This remains a first-class strategic threat.
The v1 response is not to claim that the threat disappears, but to ensure that
its effect is confined chiefly to certification latency rather than to
canonical-order steering.

\paragraph{Frontier steering.}
Because frontier state can be noisy and locally unstable, the v1 protocol does
not let $\Cfr(B)$ serve as a mandatory Grade~2 certificate gate.
This reduces the extent to which local frontier manipulation can become a
brittle strong-confirmation trigger.

\paragraph{$\Rcls$ gaming.}
If $\Rcls$ were made consensus-critical, miners could attempt to shape closure
fragility metrics so as to steer order selection itself.
PHASMDAG-Core therefore excludes $\Rcls$ from the v1 canonical-order rule and
retains it only as a diagnostic or simulation object.

These attacks cannot presently be eliminated.
The design objective is instead to prevent them from maturing into safety
failures or canonical-order steering attacks.
That is the principal sense in which PHASMDAG-Core claims robustness under
CBM pressure.

\subsection{Comparison with GhostDAG under LP Attack}

\begin{table}[t]
\centering
\caption{Preliminary security behaviour comparison under LP attack}
\label{tab:comparison}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabularx}{0.96\textwidth}{@{}L{3.2cm}YY@{}}
\toprule
\textbf{Property} & \textbf{GhostDAG} & \textbf{PHASMDAG-Core} \\
\midrule
Tentative-order basis
    & structural order also carries confirmation semantics
    & same GhostDAG-style order core \\
Strong-finality basis
    & structural order is used as proxy
    & $\Ostr(B) \wedge \mathrm{ValidCert}_{\mathcal{C}}(B)$ \\
Effect of predecessor withholding
    & may silently degrade both order and confirmation semantics
    & primarily delays Grade~2 certification \\
Effect of frontier steering
    & entangled with confirmation semantics
    & mostly pushed to Grade~3 or diagnostics \\
Consensus-steering payoff from missing closure
    & implicit and potentially large
    & deliberately bounded by design \\
\bottomrule
\end{tabularx}
\end{table}

\Cref{tab:comparison} highlights the central asymmetry:
PHASMDAG-Core does not claim to eliminate LP/BJ stress.
It instead aims to make the semantic consequences of such stress explicit by
withholding strong finality until closure has been certified.

\subsection{Certificate Coherence Theorem Target}

Under LP attack, different honest nodes may still see different local evidence
for a period of time.
The central theorem target is therefore no longer merely closure convergence,
but certificate coherence across honest nodes.

\begin{conjecture}[Certificate coherence theorem target]\label{conj:coherence}
Under CBM with honest majority, if some honest node $v$ issues
${\Certcls}_v(B)=1$, then within bounded additional time
$\Delta_{\mathrm{coherence}}$ all honest nodes will either also issue
${\Certcls}(B)=1$ or reject certification under the same deterministic
support-window evidence rules.
\end{conjecture}

\noindent If established, \Cref{conj:coherence} would show that the Grade~2
predicate is not merely a local heuristic but a consensus-meaningful object.
Without such certificate coherence, PHASMDAG-Core would remain only a
descriptively attractive semantics rather than a defensible finality rule.

\subsection{Heuristic Note on Confirmation Growth}

Any quantitative dominance claim should be treated as heuristic rather than as
the primary security result of the paper.
The primary safety improvement in PHASMDAG-Core is semantic:
strong confirmation is withheld until dependency closure is certified.

\begin{remark}[Optional analytical intuition]\label{rem:gamma_heuristic}
It remains heuristically useful to distinguish tentative-order growth from
Grade~2 confirmation growth.
Tentative order may continue to inherit the GhostDAG-style expression
$g^{(1)}_{\mathrm{PHASM}}=(1-\sigma)\gamma f$, while Grade~2 growth may be
viewed as scaled by a certificate-issuance factor $p_{\mathrm{cert}}(T)$.
However, this intuition is not the principal safety claim of the protocol and
should be understood as an analytical convenience only.
\end{remark}

% =====================================================================
\section{Formal Modelling Approach}\label{sec:formal}

This section develops a compositional formal model aligned with the three
semantic layers.

\subsection{Compositional State Machine}

Each block $B$ maintains three logical state components:

\begin{enumerate}[nosep]
    \item \textbf{Structural state} $\Sigma_{\mathrm{ord}}(B)$:
    parent set, selected-parent relation, blue-work score, ordering rank.
    This is essentially the existing GhostDAG state.

    \item \textbf{Closure state} $\Sigma_{\mathrm{cls}}(B)$:
    relevant support window $\mathcal{W}_H(B)$, the three sub-metrics
    $\Cdep(B)$, $\Ccoh(B)$, $\Cfr(B)$, unresolved-support set
    $\mathcal{U}(B)$, support-window work evidence $\mathrm{Work}_H(B)$,
    certificate-side metrics $\Cdepcert(B)$ and $\Ccohcert(B)$,
    raw closure observation $\Clvl(B)$, and certificate status $\Certcls(B)$.

    \item \textbf{Finality state} $\Sigma_{\mathrm{fin}}(B)$:
    tentative-order status, closure-certified strong-confirmation status, deep
    settlement status, and finality grade $\Ffin(B)$.
\end{enumerate}

The full state machine is the product
$\Sigma(B) = \Sigma_{\mathrm{ord}}(B) \times
\Sigma_{\mathrm{cls}}(B) \times \Sigma_{\mathrm{fin}}(B)$,
with admissible transitions defined by the protocol rules.
The transition order is strict: $\Sigma_{\mathrm{ord}}$ is computed first
(since closure depends on structural state), then $\Sigma_{\mathrm{cls}}$,
then $\Sigma_{\mathrm{fin}}$.

\subsection{Local Views and Disagreement Metrics}

The key modelling difficulty is that PHASMDAG-Core is evaluated under partial and
heterogeneous visibility.
The protocol therefore requires an explicit notation for node-local views and
for cross-node disagreement in closure scores.

\begin{definition}[Node-Local DAG View]\label{def:local_view}
Let $\Dag_v(t)$ denote the accepted header-DAG observed by honest node $v$ at
time $t$.
All protocol quantities are node-local during execution:
\[
\Ostr_v(B,t), \qquad \Clvl_v(B,t), \qquad {\Certcls}_v(B,t), \qquad \Ffin_v(B,t).
\]
Consensus safety requires that these local quantities converge sufficiently
fast after network stabilisation.
\end{definition}

\begin{definition}[Closure Disagreement]\label{def:closure_gap}
For any block $B$ visible to the honest node set $H_t$ at time $t$, define the
honest-node closure disagreement by
\begin{equation}\label{eq:closure_gap}
\Delta_{\Clvl}(B,t)
:=
\max_{u,v \in H_t}
\left|\Clvl_u(B,t) - \Clvl_v(B,t)\right|.
\end{equation}
The central modelling objective is to show that $\Delta_{\Clvl}(B,t)$ becomes
small quickly enough to support consistent Grade~2 decisions.
\end{definition}

\begin{definition}[Certificate-Coherent Strong Confirmation Target]\label{def:coherent_strong}
A block $B$ is said to satisfy the \emph{certificate-coherent strong-confirmation
target} at time $t$ if
\begin{equation}\label{eq:coherent_strong}
\forall v \in H_t,\ {\Certcls}_v(B,t)=1
\qquad\text{and}\qquad
\Delta_{\Clvl}(B,t) \le \varepsilon_{\mathrm{coh}}
\end{equation}
for a protocol tolerance $\varepsilon_{\mathrm{coh}} > 0$.
\Cref{conj:coherence} may then be read as a theorem target asserting eventual
attainment of~\eqref{eq:coherent_strong} for structurally supported honest
blocks.
\end{definition}
This convergence target is visualised in \Cref{fig:closure_convergence}.

\begin{figure}[t]
\centering
\begin{tikzpicture}[
    x=1.18cm, y=4.0cm,
    axislabel/.style={font=\scriptsize, text=black!75},
    lab/.style={font=\scriptsize, fill=white, inner sep=1.2pt, text=black!78}
]
\fill[closurecolor!7] (3.65,0.62) rectangle (5.75,0.86);
\draw[->, line width=0.9pt] (0,0) -- (6.15,0)
    node[below right, axislabel] {time $t$};
\draw[->, line width=0.9pt] (0,0) -- (0,1.02)
    node[left, axislabel] {node-local certificate-side evidence};
\draw[dashed] (0,0.62) -- (6.0,0.62);
\node[lab, anchor=east] at (0,0.62) {issuance boundary};

\draw[blueblock!80!black, thick]
    plot[smooth] coordinates {(0.35,0.22) (1.1,0.33) (2.0,0.49) (3.0,0.67) (4.2,0.78) (5.5,0.83)};
\draw[closurecolor!75!black, thick]
    plot[smooth] coordinates {(0.35,0.46) (1.1,0.52) (2.0,0.58) (3.0,0.69) (4.2,0.76) (5.5,0.81)};

\node[lab, anchor=west] at (5.56,0.83) {node $u$};
\node[lab, anchor=west] at (5.56,0.81) {node $v$};
\draw[dotted, thick] (3.6,0) -- (3.6,0.9);
\node[lab, anchor=north] at (3.6,0) {$t^\star$};
\draw[<->, semithick] (4.5,0.76) -- (4.5,0.79);
\node[lab, anchor=west] at (4.58,0.775) {cross-node gap $\le \varepsilon_{\mathrm{coh}}$};
\node[lab] at (4.7,0.69) {coherent region};
\end{tikzpicture}
\caption{Closure coherence as a convergence target.  Strong confirmation should
only become globally meaningful once node-local certificate-side evidence lies
near the issuance boundary and cross-node disagreement is sufficiently small.}
\label{fig:closure_convergence}
\end{figure}

\subsection{Block Processing Algorithm}

\Cref{alg:process} specifies the high-level block processing pipeline for
PHASMDAG-Core, making the three-layer computation order explicit.

\begin{algorithm}[t]
\caption{PHASMDAG-Core Block Processing}\label{alg:process}
\begin{algorithmic}[1]
\Require Incoming block $B_{\mathrm{new}}$ with parent set $P$
\Ensure Updated DAG state $\Sigma$
\Statex
\State \textbf{// --- Layer A: Order ---}
\State Validate header in isolation (PoW, timestamp, format) \Comment{Same as GhostDAG}
\State Check all parents exist in $\Sigma_{\mathrm{ord}}$ \Comment{\texttt{check\_parents\_exist}}
\State Compute $\Ostr(B_{\mathrm{new}})$ and $r(B_{\mathrm{new}})$ via blue-work ordering
\State Select preferred parent, compute mergeset, colour blue/red
\State Update $\Sigma_{\mathrm{ord}}(B_{\mathrm{new}})$
\Statex
\State \textbf{// --- Layer B: Closure ---}
\State Identify affected ancestors $A$ whose support windows include $B_{\mathrm{new}}$
\For{each affected block $A$}
    \State Update $\mathcal{W}_H(A)$ and $\mathcal{U}(A)$
    \State Recompute $\Cdep(A)$, $\Ccoh(A)$, and $\Cfr(A)$ \Comment{Approach A / $v_1$}
    \State $\Clvl(A) \gets g_{v_1}\!\bigl(\Cdep(A), \Ccoh(A), \Cfr(A)\bigr)$
    \State Compute $\mathrm{Work}_H(A)$
    \State Recompute $\Cdepcert(A)$ and $\Ccohcert(A)$
    \If{$\Cdepcert(A)\ge\theta_{\mathrm{dep}} \wedge \Ccohcert(A)\ge\theta_{\mathrm{coh}} \wedge \mathrm{Work}_H(A)\ge\theta_{\mathrm{work}}$}
        \State Set $\Certcls(A)\gets 1$ \Comment{Monotonic once issued}
    \EndIf
\EndFor
\State Initialise $\Sigma_{\mathrm{cls}}(B_{\mathrm{new}})$
\Statex
\State \textbf{// --- Layer C: Finality ---}
\State Compute Grade predicates using \Cref{eq:grade1,eq:grade2,eq:grade3}
\State Assign confirmation grade:
\If{$\mathsf{Grade}_3(B)$}
    \State Grade 3: Deep Settlement
\ElsIf{$\mathsf{Grade}_2(B)$}
    \State Grade 2: Strong Confirmation
\ElsIf{$\mathsf{Grade}_1(B)$}
    \State Grade 1: Tentative Ordering
\Else
    \State No confirmation
\EndIf
\State Update $\Sigma_{\mathrm{fin}}(B_{\mathrm{new}})$ and affected blocks
\end{algorithmic}
\end{algorithm}

\subsection{Raw Closure Observation and Certificate Computation}

The most critical modelling challenge is the computation of raw closure
observations together with the monotonic closure certificate.
PHASMDAG-Core treats this as a deliberately conservative design space:
one deployable certificate-oriented instantiation is standardised for $v_1$,
while more aggressive closure-aware ordering ideas are kept outside the core
protocol.

\paragraph{Approach A: Deterministic closure certification over a bounded support window ($v_1$).}
Approach~A is the \textbf{only} standardised closure computation in
PHASMDAG-Core-$v_1$.
Its simplest descendant-attestation form is the PoW-weighted statistic
\begin{equation}\label{eq:closure_attestation_work}
\Clvl_A(B) =
\frac{\sum_{B' \in \mathrm{desc}(B),\;\mathbf{1}_{\mathrm{closed}}(B')=1} W(B')}
{\sum_{B' \in \mathrm{desc}(B)} W(B')}.
\end{equation}
Here $W(B')$ denotes the PoW weight, accumulated work, or
difficulty-equivalent contribution of descendant block $B'$.
This makes closure evidence an economic-security quantity rather than a raw
descendant count, and prevents a minority miner from cheaply inflating
closure with low-value descendants alone.
It instantiates $\Cdep(B)$, $\Ccoh(B)$, and $\Cfr(B)$ exactly as in
\Cref{eq:cdep,eq:ccoh,eq:cfr}; only the first two enter the Grade~2
certificate rule in $v_1$.
Equation~\eqref{eq:closure_attestation_work} is the unwindowed intuition;
PHASMDAG-Core-$v_1$ implements the same idea through the bounded support-window
version of \Cref{eq:cdep,eq:ccoh,eq:work_window}.
It uses only structural evidence observable in the accepted header-DAG plus the
validation state required for $\mathbf{1}_{\mathrm{closed}}$.
Certificate issuance is then determined by \Cref{eq:valid_cert}.
No wall-clock propagation delay, dynamic parameter, or frontier-only trigger is
required for Grade~2 in this conservative instantiation.
This makes the construction deterministic, decomposable, auditable, and closer
to a plausible deployable core.

\paragraph{Approach B: Experimental closure-weighted ordering extension.}
An alternative line of design would feed a closure-fragility functional
directly back into ordering, for example through a score of the form
\begin{equation}\label{eq:closure_ordering_future}
\Phifn(S) = \Uord(S) - \lambda\,\Rcls(S).
\end{equation}
Such an approach is attractive because it appears to internalise LP/BJ risk
inside the ordering rule itself.
However, it also turns missing closure evidence into a strategic resource that
may be used to steer order.
For that reason, PHASMDAG-Core excludes this idea from $v_1$ and treats it as
future work.

\begin{table}[t]
\centering
\caption{Core vs.\ experimental closure-handling strategies}
\label{tab:closure_approaches}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Property} & \textbf{Approach A} & \textbf{Approach B} \\
\midrule
Deterministic across nodes & Strong & Weak--Medium \\
Ordering stability & Strong & Lower \\
Formal proof friendliness & High & Low \\
Topology bias risk & Low & Higher \\
Consensus-steering risk & Lower & Higher \\
\bottomrule
\end{tabular}
\end{table}

\begin{remark}[Version policy for closure computation]\label{rem:closure_approaches}
Approach~A is the reference instantiation for PHASMDAG-Core-$v_1$.
Approach~B remains explicitly experimental.
This is a versioning decision, not merely a modelling preference:
the base protocol optimises for semantic separation, determinism, and bounded
manipulation payoff before attempting closure-aware ordering.
\end{remark}

\subsection{Parameter Space}

The core protocol introduces the following parameters that must be analysed formally:

\begin{table}[tbp]
\centering
\caption{Protocol parameter space}
\label{tab:params}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabularx}{0.98\textwidth}{@{}L{2.7cm}L{2.6cm}Y@{}}
\toprule
\textbf{Parameter} & \textbf{Type} & \textbf{Constraints} \\
\midrule
$K$ & GhostDAG order parameter & Fixed consensus constant inherited from the
GhostDAG-style ordering core \\
$H$ & Closure window & Support-horizon size for $\mathcal{W}_H(B)$;
must be large enough to capture relevant descendant evidence \\
$\eta_{\mathrm{dep}}, \eta_{\mathrm{coh}}, \eta_{\mathrm{fr}}$ & Aggregation &
Non-negative and sum to $1$ in \Cref{eq:closure_level_v1}; diagnostic only \\
$\tau_{\mathrm{ord}}$ & Order threshold & Minimum structural support for Grade~1 \\
$\theta_{\mathrm{dep}}, \theta_{\mathrm{coh}}$ & Certificate thresholds &
Minimum dependency-completeness and coherent-support evidence for
$\mathrm{ValidCert}_{\mathcal{C}}(B)$ \\
$\theta_{\mathrm{work}}$ & Work threshold & Minimum support-window work needed
before a certificate may be issued \\
$\tau_{\mathrm{settle}}$ & Settlement threshold & Minimum deep-settlement
condition for Grade~3 \\
$\theta^{\mathrm{deep}}_{\mathrm{fr}}$ & Optional frontier stability & May be
used as a conservative Grade~3 strengthening, not as a Grade~2 gate \\
$\varepsilon_{\mathrm{coh}}$ & Coherence tolerance & Maximum admissible
cross-node disagreement when Grade~2 becomes globally meaningful \\
$\rho_{\mathrm{tent}}$ & Tentative ordering & Maximum permitted rank churn for
Grade~1 (\Cref{inv:tentative}) \\
$\varepsilon_{\mathrm{settle}}$ & Settlement target & Maximum rollback tail for
Grade~3 (\Cref{inv:settlement}) \\
$\lambda$ and $\Rcls$ & Experimental extension & Not consensus-critical in
$v_1$; retained only for diagnostics, simulation, or future variants \\
$\Clvl$ definition & Protocol design & $v_1$: Approach~A only; Approach~B is
experimental (\Cref{rem:closure_approaches}) \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Formal Verification Targets}

The following properties are targets for formal verification:

\begin{enumerate}[nosep]
    \item \textbf{Tentative bounded churn}: prove or bound
    \Cref{inv:tentative} under the blue-work order core.
    \item \textbf{Strong-confirmation safety}: prove
    \Cref{inv:strong}, namely that no two honest nodes assign conflicting
    Grade~2 confirmations at the same virtual position.
    \item \textbf{Settlement tail}: prove or upper-bound
    \Cref{inv:settlement}.
    \item \textbf{Liveness}: every honestly mined block eventually reaches
    tentative ordering (Grade~1) if the network eventually stabilises.
    \item \textbf{Certificate coherence}: establish or refute
    \Cref{conj:coherence}, which is the central theorem target of the protocol.
    \item \textbf{Payoff-bounding validation}: determine whether the bounded
    manipulation-payoff framing in \Cref{prop:graceful} survives a fuller CBM
    analysis or only a simulation-level approximation.
\end{enumerate}

% =====================================================================
\section{Engineering Verification}\label{sec:verification}

\subsection{Simulation Methodology}

The simulation programme is based on an extension of the
SimBlock~\cite{aoki2019simblock} framework to model
CBM conditions:

\begin{enumerate}[nosep]
    \item \textbf{Network model}: replace the uniform delay model with CBM,
    incorporating bandwidth limits and per-block propagation delays.
    \item \textbf{LP attack simulation}: implement the $s$-group partition
    attack from~\cite{wu2024security}, allowing the attacker to selectively
    delay block delivery to specific node groups.
    \item \textbf{Bandwidth congestion}: simulate Block Jam by enforcing a
    global bandwidth cap and measuring propagation delays as a function of
    utilisation.
    \item \textbf{Closure tracking}: augment each simulated node with the
    $v_1$ raw-closure and certificate computation (Approach~A), tracking both
    $\Clvl(B)$ and the issuance event $\Certcls(B)=1$ over time.
\end{enumerate}

\subsection{Evaluation Metrics}

The evaluation is organised around the following metrics, following the common-metrics
framework of~\cite{zhang2019common}:

\begin{table}[tbp]
\centering
\caption{Evaluation metrics}
\label{tab:metrics}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabularx}{0.94\textwidth}{@{}L{4.1cm}Y@{}}
\toprule
\textbf{Metric} & \textbf{Description} \\
\midrule
Chain growth rate $g$
    & Blocks per second entering the canonical order \\
Certificate issuance rate $p_{\mathrm{cert}}(T)$
    & Fraction of tentatively ordered blocks that become certified within horizon $T$ \\
Realised TPS
    & Actual transaction throughput under imperfect networks \\
Tentative confirmation latency
    & Time from block creation to Grade~1 \\
Strong confirmation latency
    & Time from block creation to Grade~2 \\
Deep settlement latency
    & Time from block creation to Grade~3 \\
Certificate coherence time
    & Time until honest nodes agree sufficiently for $\Certcls(B)=1$ to become globally meaningful \\
P95/P99 finality tail
    & Tail behaviour of confirmation latency under stress \\
\bottomrule
\end{tabularx}
\end{table}

\subsection{Comparative Evaluation Plan}

\begin{enumerate}[nosep]
    \item \textbf{Baseline}: GhostDAG under UDBM (optimistic reference).
    \item \textbf{GhostDAG under CBM}: GhostDAG with LP attack and bandwidth
    congestion, no mitigations.
    \item \textbf{GhostDAG + engineering mitigations}: GhostDAG with
    header-first, dependency-aware relay, and telemetry.
    \item \textbf{PHASMDAG-Core under CBM}: PHASMDAG-Core with $\Clvl$ computation
    (Approach~A) and certificate issuance, under the same CBM conditions.
\end{enumerate}

The comparison between configurations~(3) and~(4) directly tests
\Cref{prop:engineering_ceiling}: does the protocol-level closure awareness
of PHASMDAG-Core provide security benefits beyond what engineering mitigations
can achieve?

\subsection{Engineering Priorities for v1}

The conservative presentation of PHASMDAG-Core is intended to make the
protocol look buildable rather than merely suggestive.
The minimal engineering package for any $v_1$ prototype should include:
\begin{enumerate}[nosep]
    \item \textbf{Header-first propagation}, so that ordering can continue even
    when full dependency availability lags.
    \item \textbf{Dependency-aware relay}, so that blocks resolving the largest
    missing-predecessor pockets are propagated first.
    \item \textbf{Orphan and missing-predecessor recovery}, so that unresolved
    support windows can be repaired promptly once absent predecessors are
    discovered.
    \item \textbf{Compact closure-evidence caching}, so that support-window work
    and coherent-support counters are maintained incrementally rather than
    recomputed from scratch.
    \item \textbf{A fixed support horizon $H$}, so that both analysis and
    implementation remain bounded.
\end{enumerate}

\subsection{Complexity and Incremental Computation}\label{sec:complexity}

The introduction of $\Sigma_{\mathrm{cls}}$ and $\Sigma_{\mathrm{fin}}$
state machines, fixed support-window tracking, and closure-evidence caches
raises legitimate concerns about computational overhead.  A protocol that is
theoretically sound but practically unrunnable serves no one.
Incremental computation is therefore stipulated as a strict prerequisite for
$v_1$, not an afterthought.

The critical metrics that must be evaluated and bounded before a PHASMDAG-Core
prototype is considered viable are:

\begin{enumerate}[nosep]
    \item \textbf{Amortised per-block update cost for raw closure and certificate state}:
    Computing $\Clvl(B)$ for a new block requires updating the three
    sub-metrics over the affected support windows; issuing $\Certcls(B)$ also
    requires updating certificate-side thresholds over the same local region.
    Under Approach~A (\Cref{eq:cdep,eq:ccoh,eq:cfr}), this can be formulated
    incrementally: when a new block $B_{\mathrm{new}}$ arrives, only its direct
    work contribution and its canonical-child / frontier relations need to be
    propagated to affected ancestors.
    The target is $O(|\mathrm{parents}|)$ per block, not $O(|\mathrm{ancestors}|)$.

    \item \textbf{Memory footprint of closure-state indices and attestation caches}:
    Each block must store: (a) the cumulative work mass needed for
    $\Cdep(B)$, (b) support-window routing metadata needed for $\Ccoh(B)$, and
    (c) frontier concentration metadata needed for $\Cfr(B)$, together with a
    compact certificate-evidence cache for $\mathrm{Work}_H(B)$.
    These are all bounded by the finite support window $H$ and can be encoded as
    compact additive counters plus small frontier indices rather than as a full
    descendant list.

    \item \textbf{Incremental recomputation depth}:
    When $\Clvl(B)$ changes (because a new descendant arrived or a predecessor
    was resolved), raw closure scores and certificate-side margins must be
    updated along the \emph{affected descendant frontier}---not the entire DAG.
    The expected frontier size is bounded by the merge depth and the number of
    recent blocks, both of which are $O(K)$ in the current GhostDAG
    implementation.

    \item \textbf{Lazy recomputation for deep historical blocks}:
    Blocks deeper than the finality depth can be considered settled; their
    $\Clvl$ values are no longer consensus-relevant and need not be
    recomputed; certificates that have already become monotonic need not be
    re-evaluated except for audit.  Pruning strategies can treat closure and
    certificate computation similarly to how existing GhostDAG implementations
    prune anticone data beyond the pruning depth.
\end{enumerate}

\begin{remark}[Computational feasibility of Approach A]\label{rem:approach_a_cost}
Approach~A is specifically designed for incremental computation:
the work-weighted numerators and denominators in
\Cref{eq:cdep,eq:ccoh,eq:cfr} are additive or frontier-local quantities that
can be updated without re-walking the entire descendant cone.
This contrasts with the experimental ordering-aware extension
\Cref{eq:closure_ordering_future}, which would make the same quantities part of
the order-selection problem and therefore lacks a similarly clean incremental
path.
This is an additional reason to prefer Approach~A for $v_1$.
\end{remark}

\subsection{Phased Verification Roadmap}

\begin{table}[tbp]
\centering
\caption{Phased verification roadmap}
\label{tab:roadmap}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabularx}{0.98\textwidth}{@{}L{1.2cm}L{1.7cm}Y@{}}
\toprule
\textbf{Phase} & \textbf{Timeline} & \textbf{Deliverables} \\
\midrule
1 & 0--6\,mo
  & Formal model: define $\Ostr$, $\Cdep$, $\Ccoh$, $\Cfr$, $\Clvl$, $\Certcls$, and $\Ffin$;
    specify state machine and admissible transitions \\
2 & 6--12\,mo
  & Simulation: implement CBM/LP scenarios in SimBlock;
    compare GhostDAG vs PHASMDAG-Core under identical conditions \\
3 & 12--18\,mo
  & Security analysis: tentative-order growth and certificate-issuance dynamics;
    quality under heterogeneous visibility;
    adversarial threshold as function of certification degradation \\
4 & 18--24\,mo
  & Prototype: implement order layer + header-first propagation +
    dependency-aware relay + certificate issuance;
    expose three confirmation grades;
    evaluate on a controlled experimental network \\
\bottomrule
\end{tabularx}
\end{table}

Note that Phases 1--2 can proceed in parallel with the engineering mitigations
(header-first, telemetry, dependency-aware relay) and other implementation-level
studies of GhostDAG.  The CBM-GhostDAG security proof may be read as a decision
gate: if it indicates that engineering mitigations suffice at the target BPS,
PHASMDAG-Core may remain chiefly a research direction; if not, it may warrant
treatment as a more substantial redesign path.

% =====================================================================
\section{Discussion}\label{sec:discussion}

\subsection{Why Not Closure-Weighted Ordering in v1?}

Closure-weighted ordering is intellectually attractive.
It appears to internalise LP/BJ risk directly into the ordering rule itself.
However, it also turns closure evidence into a strategic resource.
If missing closure evidence can directly reduce canonical-order attractiveness,
then suppression of closure becomes a consensus-steering lever.
PHASMDAG-Core therefore adopts a deliberately conservative asymmetry:
positive closure evidence unlocks stronger finality, whereas missing closure
evidence does not directly punish canonical order.

\begin{table}[tbp]
\centering
\caption{Why PHASMDAG-Core-$v_1$ avoids closure-weighted ordering}
\label{tab:dwc_vs_closure_risk}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabularx}{0.98\textwidth}{@{}L{3.2cm}YY@{}}
\toprule
\textbf{Dimension} & \textbf{Closure-weighted ordering} & \textbf{PHASMDAG-Core-$v_1$} \\
\midrule
Role of closure
    & direct input to ordering
    & certificate-gated finality layer \\
Effect of missing evidence
    & may penalise canonical order
    & delays Grade~2 only \\
Manipulation payoff
    & may steer consensus order
    & primarily latency griefing \\
Proof friendliness
    & lower
    & higher \\
\bottomrule
\end{tabularx}
\end{table}

The choice is therefore not that closure-weighted ordering is devoid of merit.
The point is rather that it requires a separate incentive-compatibility
argument, because it makes missing closure evidence a consensus-critical
variable.
PHASMDAG-Core-$v_1$ prefers semantic separation to this more aggressive design.

\subsection{Relation to NC-Max}

The philosophical kinship between PHASMDAG-Core and NC-Max~\cite{zhang2022ncmax}
is not one of mechanism but of discipline.
NC-Max introduces a chain-style two-step confirmation process so that
confirmation does not outrun propagation and availability.
PHASMDAG-Core imports the same discipline into the DAG setting:
a robust high-throughput DAG should parallelise ordering, not finality.
Strong finality is serialized through dependency closure.

In that sense, PHASMDAG-Core should not be read as an attempt to make
GhostDAG's ordering rule ever more reactive.
It should be read as an attempt to prevent structural order from being
mistaken for strong finality once dependency availability ceases to be
well approximated by propagation completion.

\subsection{Failure Modes and Design Risks}

The proposal is best read alongside an explicit account of its likely failure
modes.  The most important of these are:
\begin{itemize}[nosep]
    \item \textbf{Certificate thresholds too conservative}: $\theta_{\mathrm{dep}}$,
    $\theta_{\mathrm{coh}}$, or $\theta_{\mathrm{work}}$ may delay Grade~2
    unduly, degrading usability.
    \item \textbf{Certificate thresholds too weak}: if the closure certificate is
    too permissive, strong confirmation may again outrun dependency maturity.
    \item \textbf{Certificate coherence slower than assumed}: if honest nodes do
    not converge quickly enough on certificate issuance, Grade~2 cannot be
    treated as a safe consensus object.
    \item \textbf{Frontier stability overused}: if $\Cfr(B)$ is allowed to creep
    back into Grade~2 semantics, local frontier steering may regain undue
    influence.
    \item \textbf{Experimental extensions misapplied}: if $\Rcls$ or $\lambda$
    are reintroduced into canonical ordering without a separate proof and
    incentive analysis, the protocol may recreate the very attack surface that
    $v_1$ was designed to avoid.
    \item \textbf{Semantic interface complexity}: the three-grade finality
    model may be theoretically well motivated yet difficult to expose through
    higher-layer interfaces without misinterpretation.
\end{itemize}

\subsection{Open Problems}\label{sec:open}

\paragraph{Certificate coherence.}
The central theorem target is no longer merely raw closure convergence.
It is certificate coherence:
whether honest nodes, given the same eventual structural evidence, converge
reliably upon the same Grade~2 decisions.
That is the hardest unresolved problem in the paper.

\paragraph{Experimental closure-weighted ordering.}
The earlier line of thought embodied by \Cref{eq:closure_ordering_future}
remains intellectually interesting.
It may prove possible to reintroduce $\Rcls$ into order selection without
creating unacceptable strategic leverage.
But that would be a separate protocol variant, not a modest modification of
PHASMDAG-Core-$v_1$.

\paragraph{$\lambda$-bearing extensions.}
In PHASMDAG-Core-$v_1$, $\lambda$ is not consensus-critical.
Any future protocol that reintroduces $\lambda$ into the core rule should use
only fixed consensus constants or hard-fork / epoch-finalised updates, and
should prove that no divergence is introduced by parameter misalignment.

\paragraph{Fork distance from GhostDAG.}
PHASMDAG-Core should perhaps not be understood as a minor variant of GhostDAG,
but rather as a more substantial protocol divergence.
If ever pursued in earnest, it would probably require a distinct protocol
identity (hence the auxiliary designation \emph{RobustDAG}) rather than a
simple suffix such as GhostDAG+.
This bears directly upon migration, client adaptation, and ecosystem
compatibility.

\paragraph{Three-grade confirmation interpretation cost.}
Moving from a single confirmation scalar to multiple semantic grades introduces
an interface problem as well as a protocol problem.
Any future deployment would require a careful theory of how higher layers
should interpret tentative ordering, strong confirmation, and deep settlement
without collapsing them back into an opaque scalar.

\subsection{Protocol-Level versus Implementation-Level Mitigation}

Implementation-level mitigations remain necessary, but they are not sufficient.
They reduce stress; they do not close the consensus-level semantic gap by
themselves.
The relevant boundary is:
\begin{itemize}[nosep]
    \item \textbf{Implementation-level mitigations can reduce} propagation skew,
    predecessor starvation, and operator blindness.
    \item \textbf{Implementation-level mitigations cannot eliminate} the hidden gap
    between header progress and strong finality while closure fragility remains
    non-consensus-visible.
    \item \textbf{Implementation-level mitigations cannot natively certify} support
    structures that are individually legal but collectively closure-fragile,
    because that requires a protocol object such as $\Certcls(B)$.
    \item \textbf{Implementation-level mitigations cannot provide} a protocol-native
    semantics for ``mature enough to trust'' without an object like
    $\Certcls(B)$, grounded in raw closure observations, entering the
    consensus mathematics.
\end{itemize}

PHASMDAG-Core still depends on such mitigations as infrastructure:
\begin{enumerate}[nosep]
    \item \textbf{Telemetry} provides the observational basis for $\Clvl(B)$
    computation and certificate diagnostics (dependency completion latency and
    orphan-age distributions).
    \item \textbf{Header-First Propagation} enables the closure layer to
    operate at the header level independently of body arrival, reducing the
    LP attack surface that the closure layer must handle.
    \item \textbf{Dependency-Aware Relay} improves the convergence rate of
    $\Clvl(B)$ by ensuring that blocks blocking the most pending successors
    are relayed first.
\end{enumerate}

The more appropriate framing is therefore: \emph{implementation-level mitigations are
foundational, but they are not a substitute for protocol redesign;
PHASMDAG-Core is the protocol-level completion that makes closure awareness explicit
in the consensus mathematics}.

% =====================================================================
\section{Conclusion}\label{sec:conclusion}

This paper has set out a preliminary formulation of PHASMDAG-Core, an
order-first, closure-certified DAG consensus protocol motivated by the security
questions raised by Wu, Wei, Zhang, and Jiang~\cite{wu2024security}.

The central contention has been that the root limitation of GhostDAG-family
protocols may lie not merely in the attackability of relay paths, but also in
the manner in which one structural object is made to stand simultaneously for
ordering, acceptability, and finality.
The proposed response is to separate structural order, dependency closure, and
finality into distinct mathematical roles.
PHASMDAG-Core is therefore not best understood as a faster GhostDAG.
It is better understood as a more semantically honest high-BPS DAG framework:
tentative ordering remains fast and parallel, whereas strong finality is made
conditional on closure certification.

Taken together, the present formulation supports the following tentative
claims:
\begin{enumerate}[nosep]
    \item Block Jam and Late-Predecessor effects can be brought within the
    protocol's mathematical core rather than treated solely as external
    implementation concerns.
    \item A concrete and auditable $v_1$ instantiation of closure can be given
    through bounded, work-weighted support-window evidence and the monotonic
    certificate $\Certcls(B)$.
    \item The main safety benefit is semantic rather than speed-theoretic:
    strong confirmation is withheld until dependency closure is certified.
    \item The protocol can be framed so that closure manipulation is not made
    to disappear, but its payoff is deliberately bounded by refusing to let it
    become a cheap canonical-order steering lever.
    \item Semantically explicit confirmation grades can be coupled to explicit
    invariants rather than collapsed into a single opaque scalar.
\end{enumerate}

The principal unresolved theorem target is certificate coherence:
whether honest nodes, once presented with sufficiently mature structural
evidence, converge reliably upon the same Grade~2 decisions.
Approach~A has been standardised as the reference $v_1$ design, while
closure-weighted ordering remains future work.
The paper has therefore offered not a completed proof, but a conservative
protocol proposal together with a phased programme of formal modelling,
simulation, and implementation.

Whether PHASMDAG-Core ultimately warrants treatment as an independent protocol
family---or, in the auxiliary language used here, as a genuine RobustDAG
design---must depend upon the outcome of that programme.

% =====================================================================
\section*{Acknowledgements}

The author thanks colleagues and interlocutors whose discussions on DAG
consensus, delayed acceptability, and finality semantics helped sharpen the
questions addressed in this paper.
Any remaining errors, omissions, or overstatements are the sole
responsibility of the author.

% =====================================================================
\bibliographystyle{plain}

\begin{thebibliography}{10}

\bibitem{sompolinsky2013secure}
Y.~Sompolinsky and A.~Zohar,
``Secure High-Rate Transaction Processing in Bitcoin,''
\emph{IACR Cryptology ePrint Archive}, Report 2013/881, 2013.

\bibitem{garay2015backbone}
J.~Garay, A.~Kiayias, and N.~Leonardos,
``The Bitcoin Backbone Protocol: Analysis and Applications,''
\emph{Advances in Cryptology -- EUROCRYPT}, 2015.

\bibitem{eyal2013majority}
I.~Eyal and E.~G. Sirer,
``Majority Is Not Enough: Bitcoin Mining Is Vulnerable,''
\emph{arXiv preprint arXiv:1311.0243}, 2013.

\bibitem{sompolinsky2016spectre}
Y.~Sompolinsky and A.~Zohar,
``SPECTRE: A Fast and Scalable Cryptocurrency Protocol,''
\emph{IACR Cryptology ePrint Archive}, Report 2016/1159, 2016.

\bibitem{sompolinsky2018phantom}
Y.~Sompolinsky and A.~Zohar,
``PHANTOM: A Scalable BlockDAG Protocol,''
\emph{IACR Cryptology ePrint Archive}, Report 2018/104, 2018.

\bibitem{sompolinsky2021ghostdag}
Y.~Sompolinsky, S.~Wyborski, and A.~Zohar,
``PHANTOM GHOSTDAG: A Scalable Generalization of Nakamoto Consensus,''
\emph{Proceedings of the 3rd ACM Conference on Advances in Financial
Technologies (AFT)}, pp.~57--70, 2021.

\bibitem{wu2024security}
S.~Wu, P.~Wei, R.~Zhang, and B.~Jiang,
``Security-Performance Tradeoff in DAG-based Proof-of-Work Blockchain
Protocols,''
\emph{Proceedings of the Network and Distributed System Security Symposium
(NDSS)}, 2024.

\bibitem{zhang2019common}
R.~Zhang and B.~Preneel,
``Lay Down the Common Metrics: Evaluating Proof-of-Work Consensus Protocols'
Security,''
\emph{Proceedings of the 40th IEEE Symposium on Security and Privacy
(S\&P)}, pp.~1441--1458, 2019.

\bibitem{yu2021ohie}
H.~Yu,
``OHIE: Blockchain Scaling Made Simple,''
\emph{Proceedings of the IEEE Symposium on Security and Privacy (S\&P)},
pp.~908--925, 2021.

\bibitem{bagaria2019prism}
V.~Bagaria, S.~Kannan, D.~Tse, G.~Fanti, and P.~Viswanath,
``Prism: Deconstructing the Blockchain to Approach Physical Limits,''
\emph{Proceedings of the ACM SIGSAC Conference on Computer and Communications
Security (CCS)}, pp.~585--602, 2019.

\bibitem{aoki2019simblock}
Y.~Aoki, K.~Shigeta, and R.~Banno,
``SimBlock: A Validated Blockchain Simulator,''
\emph{Proceedings of the IEEE International Conference on Blockchain
(Blockchain)}, pp.~131--139, 2019.

\bibitem{zhang2022ncmax}
L.~Zhang, F.~Zhang, and J.~Sun,
``NC-Max: Breaking the Security-Performance Tradeoff in Nakamoto Consensus,''
\emph{Proceedings of the Network and Distributed System Security Symposium
(NDSS)}, 2022.

\end{thebibliography}

% =====================================================================
\clearpage
\appendix
\section{Relationship to GhostDAG's \texorpdfstring{$K$}{K} Parameter}\label{app:k_param}

GhostDAG's security parameter $K$ is derived from the network delay bound $D$
and block production rate $f$ under UDBM assumptions~\cite{sompolinsky2021ghostdag}:
\begin{equation}\label{eq:k_original}
K = f(2Df,\, \delta)
\end{equation}
where $\delta$ is the tail probability that the anticone of a block exceeds
$K$.  The derivation assumes that blocks are accepted within delay $D$.

Under CBM, the effective delay is $\Delta_{\mathrm{actual}} = D + \Delta_{\mathrm{proc}}$,
which may exceed $D$.  A na\"ive adaptation would replace $D$ with
$D_{\mathrm{eff}} = D + \mathbb{E}[\Delta_{\mathrm{proc}}]$:
\begin{equation}\label{eq:k_naive}
K_{\mathrm{naive}} = f(2D_{\mathrm{eff}}f,\, \delta)
\end{equation}
However, this approach has two problems:
\begin{enumerate}[nosep]
    \item $\mathbb{E}[\Delta_{\mathrm{proc}}]$ is not a fixed constant but depends on
    network conditions and attack intensity, making $K_{\mathrm{naive}}$ unstable.
    \item Changing $K$ is a hard consensus change---all nodes must agree on the
    same $K$---but $D_{\mathrm{eff}}$ varies across nodes under LP attack.
\end{enumerate}

PHASMDAG-Core resolves this differently.  The closure layer does \emph{not} alter
the consensus value of $K$, nor does it reweight an otherwise valid ordered
block inside the order core.
Instead, effective delay exposure appears in the transition from tentative
order to strong confirmation:
\begin{itemize}[nosep]
    \item The $K$ parameter remains a consensus constant derived from the
    optimistic $D$ (as in current GhostDAG).
    \item Tentative order is still computed from the GhostDAG order core under
    that fixed $K$.
    \item The closure layer determines whether an already ordered block may
    advance from Grade~1 to Grade~2 by obtaining $\Certcls(B)=1$.
    \item Under LP attack, the observable degradation therefore appears as a
    lower certificate-issuance rate $p_{\mathrm{cert}}(T)$ rather than as a
    consensus-visible increase in $K$.
\end{itemize}

This design choice is deliberate: it keeps the consensus rule simple and
stable ($K$ is fixed) while allowing adaptive delay awareness to enter through
certificate-gated finality.  The tradeoff is that strong-finality growth may
degrade under LP attack even while tentative order continues to grow according
to the underlying GhostDAG order core.

\begin{table}[H]
\centering
\caption{Parameter handling comparison}
\label{tab:k_comparison}
\small
\renewcommand{\arraystretch}{1.1}
\begin{tabularx}{0.98\textwidth}{@{}L{3.8cm}YY@{}}
\toprule
\textbf{Approach} & \textbf{Consensus $K$} & \textbf{Delay handling} \\
\midrule
GhostDAG (UDBM) & $K = f(2Df,\delta)$ & $D$ assumed uniform \\
GhostDAG + adaptive $K$ & $K = f(2D_{\mathrm{eff}}f,\delta)$ & Consensus risk: nodes disagree on $D_{\mathrm{eff}}$ \\
PHASMDAG-Core & $K = f(2Df,\delta)$ (unchanged) & Order unchanged; delay appears in certificate issuance $\Certcls(B)$ \\
\bottomrule
\end{tabularx}
\end{table}

\end{document}
