\documentclass[a4paper,11pt]{article}

\usepackage[a4paper,margin=2.55cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,fit,shapes.geometric}

\definecolor{LernaBlue}{HTML}{205493}
\definecolor{LernaGreen}{HTML}{2C7A58}
\definecolor{LernaAmber}{HTML}{A86500}
\definecolor{LernaGrey}{HTML}{5E6672}

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

\hypersetup{
  colorlinks=true,
  linkcolor=LernaBlue,
  citecolor=LernaBlue,
  urlcolor=LernaBlue,
  pdftitle={Lerna Protocol: Hydra-native Virtual Channel Factories for Cardano eUTxO},
  pdfauthor={Arthur Zhang}
}

\lstdefinestyle{lerna}{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  columns=fullflexible,
  backgroundcolor=\color{gray!4},
  rulecolor=\color{gray!35},
  keepspaces=true
}
\lstset{style=lerna}

\newtheorem{definition}{Definition}
\newtheorem{assumption}{Assumption}
\newtheorem{invariant}{Invariant}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}

\newcommand{\Lerna}{\mathsf{Lerna}}
\newcommand{\Head}{\mathsf{Head}}
\newcommand{\Factory}{\mathsf{Factory}}
\newcommand{\VC}{\mathsf{VC}}
\newcommand{\State}{\mathsf{State}}
\newcommand{\Header}{\mathsf{Header}}
\newcommand{\Manifest}{\mathsf{Manifest}}
\newcommand{\Touched}{\mathsf{Touched}}
\newcommand{\Proof}{\mathsf{Proof}}
\newcommand{\Desc}{\mathsf{Desc}}
\newcommand{\Plan}{\mathsf{Plan}}
\newcommand{\Receipt}{\mathsf{Receipt}}
\newcommand{\hash}{\mathsf{H}}
\newcommand{\sig}{\mathsf{Sig}}
\newcommand{\ada}{\mathsf{ADA}}

\title{\textbf{Lerna Protocol}\\
\large Design Note for Hydra-native Virtual Channel Factories on Cardano eUTxO}
\author{Arthur Zhang}
\date{5 June 2026}

\begin{document}
\pagenumbering{Alph}

\begin{titlepage}
\centering
\vspace*{1.15cm}
{\sffamily\bfseries\LARGE Lerna Protocol\par}
\vspace{0.35cm}
{\sffamily\large Hydra-native Virtual Channel Factories for Cardano eUTxO\par}
\vspace{1.25cm}
{\normalsize Arthur Zhang\par}
\vspace{0.2cm}
{\normalsize 5 June 2026\par}
\vspace{1.25cm}

\begin{abstract}
\noindent
Lerna Protocol is a proposed design note for a virtual-channel factory
discipline in Hydra-family Cardano eUTxO systems. It defines a Hydra-native
method for separating stable value authority, moving signed state evidence,
partitioned conservation, touched leaves, and local proof discipline. Lerna is
not ``eltoo for Cardano'', not a replacement for Hydra, and not a Hydrozoa
competitor. Its narrower claim is that a Head-like ledger can host compact
rights and virtual channels, provided those rights remain deterministically
unfoldable into ordinary ledger outputs.

\medskip
\noindent
The layer boundary is the paper's main point. Hydra virtualises Cardano
execution inside one Head. Interhead virtualises Heads across Heads. Lerna
virtualises rights and channels inside a physical Head, an Interhead virtual
Head, or a sufficiently Head-like ledger. Hydrozoa decouples execution from
custody and specifies deterministic outer rollout; Lerna borrows that design
ethic as inner materialisation of compact factory rights into Head UTxOs before
outer fanout or rollout. Compactness is therefore not free compression. It is a
debt that must be repaid before the outer settlement mechanism is asked to
finalise concrete outputs.

\medskip
\noindent
The concrete V0 target is deliberately modest: ADA-only, Head-local, and
dependent on existing Hydra close, contest, and fanout semantics. V0 safety
requires at least one authorised Head participant, operator, or delegated
watcher to checkpoint or materialise the latest relevant Lerna state into a
Hydra snapshot before contestation expires. If that assumption is unacceptable,
the application must move to a future V1 claim-UTxO or another claim design;
this note does not claim V1 feasibility. The originality claim is limited to
three prototype obligations: C1, an inner factory-rights discipline; C2,
deterministic inner materialisation before outer fanout or rollout; and C3,
factory-local non-interference plus release receipts.
\end{abstract}

\vfill
\begin{flushleft}
\textbf{Keywords:} Cardano, Hydra, Hydrozoa, eUTxO, virtual channels, channel
factories, native assets, Aiken, Plutus, state channels, non-interference.
\end{flushleft}
\end{titlepage}

\pagenumbering{roman}
\tableofcontents
\clearpage
\pagenumbering{arabic}

\section{Introduction and Motivation}

Hydra Head is the first practical member of the Hydra family. It lets a fixed
set of participants run Cardano-like transactions off-chain with low latency,
confirm state in signed snapshots, and return the resolved UTxO set to L1
through close, contest, and fanout. Its implementation path also includes
incremental deposits, decommit-style exits, and partial fanout work. These
features make Hydra a natural substrate for application-specific L2 protocols,
but they expose a precise economic and operational question: must every
bilateral channel, balance, rebalance, and local exit become a Head-visible
UTxO while the application is still in its fast path?

Lerna answers no. It proposes a Head-local factory state that can hold many
virtual-channel rights compactly while retaining deterministic paths back to
ordinary Cardano outputs. The protocol is aimed at wallets, payment
applications, app-specific liquidity groups, and decentralised applications
that want the speed of Hydra, including Hydra Pay-like application workflows,
but do not want to represent every temporary virtual channel as a full Head
UTxO during normal operation~\cite{hydra-pay}.

The positioning is deliberately narrow: Hydra virtualises Cardano execution
inside one Head; Interhead virtualises Heads across Heads; Lerna virtualises
rights and channels inside a Head-like ledger. Compactness is useful only if it
has a repayment rule. Before outer settlement, compact rights must be unfolded
into Cardano-valid outputs, reabsorbed under fresh agreement, or deliberately
released.

This work is motivated by the need for a Cardano-native virtual-channel factory
that keeps stable funding, moving state evidence, operator publication duties,
partitioned conservation, and factory-local proofs explicit inside the Hydra
execution model. Cardano provides eUTxO, native assets, Plutus and Aiken
validators, inline datums, reference inputs, reference scripts, min-ADA,
transaction validity intervals, and Hydra's own Head state machine. Lerna's
reviewable claims therefore stand on Cardano/Hydra semantics, formal V0 objects,
and prototype evidence.

The practical problem is not only throughput. It is exit discipline. A virtual
channel factory is useful only if the participants can tell, at every point,
what would happen if the outer Head stopped making progress. Lerna therefore
treats materialisation as a first-class protocol object. Before a Head fans out,
Lerna V0 must expand every unresolved virtual right into ordinary Head UTxOs or
agreed zero-value releases. Lerna V1 may instead fan out a compact claim object,
but only if a real validator design can enforce latest-state settlement and
non-interference on L1.

\paragraph{Research contributions and claim scope.}
This paper makes three Cardano-specific design contributions. They are stated
as protocol obligations and prototype targets, not as finished cryptographic
theorems.

\begin{description}[leftmargin=2.4cm,style=nextline]
\item[C1. Inner factory rights discipline.]
Lerna sketches compact rights for virtual channels inside a physical Hydra Head
or an Interhead virtual Head. These rights are not miniature Heads and not L1
claims in V0. They are Head-local accounting claims over ADA balances,
reserves, memberships, local exits, and settlement descriptors.

\item[C2. Deterministic inner materialisation before outer fanout or rollout.]
Lerna turns compact factory rights into ordinary Head UTxOs before Hydra fanout,
or before an outer Hydrozoa-style rollout. Compactness is therefore not treated
as free compression; it creates an explicit materialisation debt that the
protocol must repay before the outer channel settles.

\item[C3. Factory-local non-interference plus release receipts.]
Lerna defines the proof obligation for reduced-signature local exits: touched
leaves and a rights-dependency graph must show that untouched participants are
not affected. The ADA-only V0 prototype should use full factory authorisation
unless this predicate is implemented and tested. Release receipts then remove
materialised rights from compact state, preventing the same right from being
claimed both as a factory entry and as a Head UTxO.
\end{description}

These contributions are not citation claims. They are protocol obligations that
can be falsified by a prototype: if C1 cannot be represented as a compact
Head-local state object, Lerna collapses into ordinary Head UTxO management; if
C2 cannot be completed within the Head's liveness envelope, compactness is
unsafe; if C3 cannot be checked without full-factory signatures in ordinary
local exits, the first prototype should honestly remain a full-authorisation
factory rather than claim reduced-signature security.

Equivalently, the originality claim is not that Lerna discovers state channels,
virtual heads, deterministic rollout, or latest-state settlement. Those ideas
already exist in the surrounding literature. The claim is that a Hydra-family
ledger needs a smaller internal discipline: compact rights that are cheap while
the Head is open, but that carry an explicit repayment rule before the outer
ledger is asked to settle concrete UTxOs. This is the point at which the three
contributions meet: C1 creates compact inner rights, C2 repays their fanout
debt, and C3 prevents that repayment from double-spending or mutating untouched
rights.

\begin{remark}[ADA-only first claim]
The only implementation-level claim made for V0 in this note is an ADA-only
Head-local factory. Native assets are important to the Cardano design space, but
they are treated here as an extension and benchmark agenda because Hydra's
current known limitations around unburned native tokens and intrinsically large
outputs can directly invalidate careless fanout plans.
\end{remark}

\section{Why Not ``Eltoo for Cardano''}

The eltoo lineage is useful but narrow. It gives one important intuition: a
newer mutually authorised state should supersede stale evidence without
penalty. That intuition is not enough to define a Cardano protocol.

Lerna is not eltoo for three reasons. First, Hydra already has a dispute
surface. A Head closes with a snapshot; participants may contest with a newer
snapshot; after the contestation deadline, fanout distributes the resolved UTxO
set. Lerna must fit into this close/contest/fanout mechanism rather than invent
a parallel L1 replacement game for V0.

Secondly, Cardano's accounting is not Bitcoin's output scripting model. Even
the ADA-only V0 path must account for lovelace, min-ADA, script datums,
reference inputs, reference scripts, and validator resource limits; a later
multi-asset profile must also account for native assets by policy ID and asset
name. A ``newer state wins'' slogan does not define how a compact factory state
materialises into valid eUTxO outputs.

Thirdly, the Cardano community already has Hydra-family research directions.
Interhead Hydra studies composition across Heads. Hydrozoa explores a flexible
multi-party state-channel design that decouples execution from custody and has
a rule-based fallback regime. Lerna is positioned as an inner virtual-channel
factory layer that can live inside these systems. It is not a replacement
narrative for any of them.

\begin{remark}[Positioning boundary]
Hydra provides the outer L2 state channel and snapshot/fanout machinery.
Hydrozoa-like systems may provide more flexible custody, dynamic membership,
and fallback regimes. Lerna provides factory-local semantics for virtual
channels, latest-state headers, release receipts, non-interference proofs, and
materialisation plans.
\end{remark}

\section{Related Work}

\subsection{Hydra Head}

Hydra Head is an isomorphic multi-party state channel for Cardano~\cite{hydra-paper}.
Participants deposit UTxOs, run Cardano-like transactions off-chain, confirm
state in snapshots, close with a snapshot if needed, contest stale snapshots,
and fan out the resolved state back to L1~\cite{hydra-docs}. The implementation and
documentation are directly relevant for Lerna because they define the outer
lifecycle and the practical limits that Lerna must respect: participant limits,
transaction size and execution-budget limits, deposit and decommit mechanics,
snapshot numbering, contestation deadlines, and partial fanout~\cite{hydra-repo}.

The known limitations are especially important. Hydra documentation warns that
transaction-size and execution-budget limits bound the number of participants
and the final UTxO set that can be fanned out, and it still highlights two
application-facing hazards: too many Head UTxOs can prevent finalisation, and
tokens minted but not burned inside an open Head can also prevent
finalisation~\cite{hydra-known}. Partial fanout and improved fanout costing can
reduce batching pressure, but they do not make an intrinsically invalid,
over-sized, or badly tokenised output safe. Lerna's compact factory state
directly responds to this pressure: it avoids making every virtual channel a
Head-visible UTxO during the fast path. The response is only safe because
compact rights later undergo deterministic materialisation into valid Head
UTxOs before outer settlement.

The closing consequence for Lerna is precise: Hydra is the outer settlement and
fanout substrate, and Lerna V0 does not modify that substrate. Lerna behaves as
a Head-local ADA application whose compact state is unfolded before fanout.
Later profiles that use native assets, large inline datums, or mint/burn
boundaries must prove compatibility with Hydra finalisation separately.

\subsection{Interhead Hydra}

Interhead Hydra studies virtual Hydra heads across two partially overlapping
Hydra Heads~\cite{interhead}. Its construction operates an Interhead state
machine in two Heads, lets intermediaries who belong to both Heads provide
collateral, and can convert the virtual Head into a regular Hydra Head if the
optimistic path fails. The paper is especially important for Lerna because it
establishes a Cardano-native way to think about virtual Head-like state without
opening that Head on L1 in the optimistic case.

The closing distinction is the layer boundary. Interhead virtualises a Hydra
Head across Heads. Lerna virtualises rights and channels inside a physical Head
or inside an Interhead virtual Head. Interhead's main hard problem is
synchronising partial state machines and collateralising a virtual Head across
intermediaries. Lerna's main hard problem is proving factory-local
non-interference and completing deterministic materialisation before the
containing Head-like ledger settles.

\subsection{Hydrozoa}

Hydrozoa is a Cardano multi-party state-channel project that explicitly
separates execution from custody~\cite{hydrozoa-spec}. Its draft specification
uses a multisig regime for the optimistic case, a rule-based Plutus regime for
fallback, an L2 ledger of active UTxOs, minor/major/final blocks, deposits,
withdrawals, treasury state, dispute votes, and deterministic rollout of
withdrawn UTxOs that do not fit in one Cardano transaction. These choices are
highly relevant to Lerna because they show a mature Cardano-specific design
pattern: ordinary operation may be cheap and multisig-like, but the exit path
must be deterministic, bounded by Cardano transaction limits, and explicit about
which state version is being settled.

Lerna should use Hydrozoa cautiously. The local repository notes that Hydrozoa
is pre-alpha and that public specifications are in flux~\cite{hydrozoa}. Even
so, three architectural lessons are stable enough to use as design constraints:
execution and custody should be separated; deposits and withdrawals need
explicit deadlines and margins; and large output sets need deterministic
rollout rather than ad hoc transaction construction. The closing consequence is
not competition: Hydrozoa studies outer rollout and execution-custody
separation, while Lerna borrows that design ethic as inner rollout and
materialisation of compact factory rights.

\subsection{Eltoo, Channel Factories, and Perun}

Eltoo is relevant as an intellectual ancestor for non-punitive latest-state
settlement~\cite{eltoo}. Lerna adopts monotonic state numbers and rejects
penalty-based settlement as the default. It does not adopt Bitcoin-specific
sighash assumptions, transaction replacement mechanisms, or output templates.

Channel factories show why shared funding can support many off-chain
relationships~\cite{factories}. Perun and related virtual-channel work show
the broader adjudication pattern: off-chain state becomes enforceable through a
dispute or adjudication mechanism~\cite{perun}. The closing boundary is that
these works provide intellectual lineage only. Lerna does not take
Bitcoin-specific assumptions, and it reserves heavyweight Cardano adjudication
for the future V1 claim-UTxO mode, whose feasibility is not claimed here.

\section{System Model}

\subsection{Cardano L1}

Cardano L1 is the settlement layer. It supplies eUTxO accounting, native
assets, Plutus and Aiken validators, reference inputs, inline datums, reference
scripts, transaction validity intervals, and min-ADA requirements. Lerna
assumes no Cardano consensus change~\cite{cardano-eutxo}.

\subsection{Hydra Head}

A Hydra Head is the outer L2 execution environment. Participants submit
transactions to Hydra nodes; the Head state evolves through signed snapshots.
Any participant can close with a snapshot; others may contest with newer valid
snapshots before the contestation deadline; final fanout distributes the
resolved Head UTxO set back to L1. Lerna V0 assumes an open Head and uses only
this existing machinery.

\subsection{Lerna Factory}

A Lerna factory is a compact Head-local object representing many virtual
channels and their rights. In V0 it is a Head-internal UTxO or state object
controlled by a validator or by agreed Head-local transaction rules. It
contains commitments to balances, subchannels, membership, reserve claims,
release receipts, and materialisation descriptors.

\subsection{Virtual Participants, Watchers, and Operators}

Virtual-channel participants may be the same as Head participants, a subset of
them, or users represented by Head participants. The distinction matters. If a
virtual participant is not a Head participant, the design must state who can
submit, observe, contest, or materialise on that participant's behalf. Lerna V0
therefore treats operator and watcher responsibilities as part of the
deployment profile.

Hydra already requires participants to remain sufficiently online for safe
contestation. Lerna adds inner obligations: tracking latest virtual-channel
states, checking release receipts, detecting stale local exits, and preparing
materialisation plans before Head closure. Operators may pay ordinary
Head-level transaction fees or manage application liquidity. They do not
receive implicit authority over factory-owned balances.

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  box/.style={draw, rounded corners=2pt, align=center, text width=3.25cm,
              minimum height=0.7cm, inner sep=5pt},
  arrow/.style={-Latex, thick, draw=LernaGrey}
]
\node[box, fill=LernaBlue!7, draw=LernaBlue] (l1) {Cardano L1\\eUTxO, validators, native assets};
\node[box, fill=LernaGreen!8, draw=LernaGreen, below=0.55cm of l1] (hydra) {Hydra-family Head\\snapshots, close, contest, fanout};
\node[box, fill=LernaAmber!12, draw=LernaAmber, below=0.55cm of hydra] (factory) {Lerna factory\\compact roots and materialisation plan};
\node[box, fill=gray!8, draw=gray!70, below left=0.55cm and -0.15cm of factory] (vc1) {Virtual channel\\latest-state updates};
\node[box, fill=gray!8, draw=gray!70, below right=0.55cm and -0.15cm of factory] (vc2) {Local exits\\release receipts};
\draw[arrow] (l1) -- (hydra);
\draw[arrow] (hydra) -- (factory);
\draw[arrow] (factory) -- (vc1);
\draw[arrow] (factory) -- (vc2);
\end{tikzpicture}
\caption{Lerna as an inner factory layer. Cardano L1 and Hydra supply the
outer settlement and state-channel machinery; Lerna supplies compact
factory-local semantics inside the Head.}
\end{figure}

\section{Threat Model and V0 Enforcement Assumptions}

Lerna V0 is conservative because it avoids a new L1 dispute protocol. That
choice has a precise cost: a virtual participant does not receive an independent
L1 exit path merely by participating in a Lerna virtual channel. In V0, the
enforceable outer object is still the Hydra snapshot that survives close and
contest. V0 safety therefore requires a concrete liveness path from the latest
Lerna evidence to a Hydra snapshot before the contestation window closes.

\subsection{Actors and Authority}

\begin{description}[leftmargin=3.1cm,style=nextline]
\item[Head participant.]
A Hydra party that can run a Hydra node, sign snapshots, submit Head-local
transactions, close the Head, and contest with a newer snapshot.

\item[Virtual participant.]
A user or application endpoint whose right is represented inside the Lerna
factory. A virtual participant may also be a Head participant, but V0 does not
assume this.

\item[Operator.]
An application actor that may batch updates, fund ordinary transaction fees, or
coordinate materialisation. An operator has no implicit authority to move
participant balances unless the relevant transition signature set authorises
that movement.

\item[Watcher.]
An actor that observes Lerna latest-state evidence and Hydra close events. In
V0 a watcher is useful only if it has a path to an authorised Head participant
or is itself a Head participant. Observation without submission authority is
not an enforcement path.
\end{description}

\subsection{Adversary}

The adversary may corrupt any subset of virtual participants, operators,
watchers, and Head participants. It may withhold signatures, withhold proof
material after signing a header, propose same-number conflicting states, refuse
to include a valid Lerna checkpoint in a Head snapshot, close a Head with a
stale snapshot, delay materialisation until the remaining contestation window
is too short, and create materialisation pressure by opening many compact
rights. It may also attempt ordinary replay across networks, Heads, factories,
epochs, descriptor versions, and signature domains.

V0 does not assume an honest majority of Head participants. Its liveness
condition is weaker and sharper: for every disputed right there must exist at
least one non-corrupt actor with both the evidence and the authority path needed
to get the newer Lerna checkpoint or materialised outputs into a newer Hydra
snapshot before the contestation deadline. If every Head participant able to
move the state forward is corrupt, censored, or offline, V0 offers no separate
Lerna-level remedy.

\subsection{Censorship Boundary}

The Head-internal censorship attack is fundamental. A non-Head virtual
participant may know the latest valid virtual state and still be unable to
force it into a Hydra snapshot. In V0 that user's practical protection is
delegation to a Head participant, an operator who is also a Head participant,
or a watcher with a live submission path. This makes V0 a cooperative or
watcher-assisted Head-local accounting layer. It is not a unilateral-exit
protocol for arbitrary non-Head users.

\paragraph{Sharp V0 assumption block.}
V0 safety requires at least one authorised Head participant, operator, or
delegated watcher to checkpoint or materialise the latest relevant Lerna state
into a Hydra snapshot before contestation expires. This is not an availability
optimisation; it is the enforcement bridge between compact inner rights and the
outer close/contest/fanout substrate. If a deployment cannot make this bridge
credible, V0 is the wrong mode and the design must move to V1 claim-UTxO
research, with the validator feasibility burden made explicit.

\begin{assumption}[Checkpoint-or-materialise liveness]
For every latest Lerna state that may affect final rights, at least one
authorised Head participant, operator, or delegated watcher can checkpoint that
state or materialise its affected rights into the Hydra Head, and can ensure
that the resulting Hydra snapshot is available before any relevant contestation
deadline expires.
\end{assumption}

\begin{assumption}[Contestability]
If a stale Hydra snapshot is used to close the Head, at least one honest
watcher can observe the stale close, construct or obtain a newer Hydra snapshot
containing the latest Lerna checkpoint or materialisation, and contest before
the Hydra contestation deadline. The watcher may be a Head participant, an
operator who is also a Head participant, a delegated watcher coordinating with
a Head participant, or an application-specific watching service with an
explicit submission path. V0 does not give a non-Head virtual participant an
independent L1 claim.
\end{assumption}

\paragraph{Who can act?}
The actor that moves latest Lerna evidence into a Hydra snapshot must be
authorised at the Head boundary. In a minimal deployment this actor is a Head
participant. In an application deployment it may be an operator who is also a
Head participant, a delegated watcher coordinating with a Head participant, or
an application-specific watching service with a clearly specified submission
path. Observation alone is not enough: the deployment must identify the key,
node, or service that can actually cause the newer snapshot or materialisation
transaction to be proposed and signed. V0 does not specify a general delegation
protocol for non-Head virtual participants; it requires the deployment to make
that delegation explicit.

\begin{assumption}[Materialisation capacity]
The materialisation workload induced by the current factory state fits inside
the deployment's transaction-size, execution-budget, min-ADA, native-asset, and
contestation-time envelope. If this cannot be maintained with margin, the
factory must enter shutdown-preparation mode and stop accepting new compact
rights.
\end{assumption}

\begin{remark}[When V0 is not enough]
If these assumptions are unacceptable for a target application, the application
must move to V1 or another claim-UTxO design. V0 is a practical Hydra-local
factory discipline, not a universal unilateral-exit protocol for users who are
entirely outside the Head participant set. This paper does not present V1 as
deployable.
\end{remark}

\begin{remark}[Delegation open question]
The unresolved V0 engineering question is how to encode watcher or operator
delegation without making the operator a custodian of factory balances. A
prototype should report the exact actor that can submit each checkpoint,
materialisation, and contest transaction.
\end{remark}

\section{Design Thesis}

Hydra provides fast isomorphic eUTxO execution. Lerna provides virtual-channel
factory discipline inside Hydra-family Heads.

The thesis has five parts.

\begin{enumerate}[label=\textbf{T\arabic*.},leftmargin=2.6em]
\item \textbf{Compactness.} Many virtual channels can be represented by one
factory state object inside a Head, avoiding Head-internal UTxO explosion
during normal operation.
\item \textbf{Deterministic unfoldability.} Compact state must always admit a
deterministic materialisation plan into ordinary Head UTxOs, because Hydra
fanout ultimately distributes UTxOs rather than abstract rights.
\item \textbf{Latest-state settlement.} Each virtual channel and factory
transition uses monotonic state numbers and domain-separated signatures. Newer
valid Lerna states supersede older Lerna states at the factory layer, without
penalties.
\item \textbf{Non-interference.} Reduced signing sets are safe only when
touched leaves and dependency proofs establish that untouched participants'
rights are unchanged.
\item \textbf{V0 before V1.} The first practical implementation should keep
all Lerna enforcement inside an open Hydra Head. L1 claim UTxOs are future
research unless and until validators can enforce the necessary semantics within
Cardano limits.
\end{enumerate}

\section{Lerna below Interhead: from virtual Heads to virtual rights}

Interhead Hydra virtualises a Hydra Head across two partially overlapping
Heads~\cite{interhead}. Its endpoints want something Head-like: an execution
environment whose optimistic path can stay off-chain, whose intermediaries can
provide collateral across the boundary, and whose failure path can convert the
virtual Head into regular enforceable Hydra state. The object being
virtualised is therefore a Head.

Lerna virtualises a smaller object. It does not create a virtual Hydra Head,
does not add a cross-Head state machine, and does not inherit Interhead's
conversion argument. A Lerna virtual channel is a right-bearing relationship
inside a factory: a compact claim over balances, reserves, membership, local
exit paths, and settlement descriptors. Its natural home is one physical Hydra
Head, but the same factory state can in principle live inside an Interhead
virtual Head. In that composition, Interhead supplies the virtual execution
container and Lerna supplies the inner rights discipline.

The distinction can be summarised as:
\[
\begin{array}{lll}
\text{Hydra Head} &:& \text{virtualises Cardano execution inside one Head},\\
\text{Interhead} &:& \text{virtualises a Head across Heads},\\
\text{Lerna} &:& \text{virtualises rights and channels inside a Head-like ledger}.
\end{array}
\]

This positioning changes the proof obligations. Interhead must synchronise
partial state machines, intermediary collateral, and conversion phases. Lerna
must show factory-local non-interference and deterministic materialisation.
The two are compatible, but neither proof discharges the other. If a Lerna
factory lives inside an Interhead virtual Head, timeout alignment becomes a
first-class parameter: the inner materialisation schedule must complete early
enough for the outer Interhead escalation and the eventual Hydra close,
contest, and fanout path.

The interface between the two layers should therefore be narrow. Interhead need
only see the Lerna factory as part of the virtual Head ledger state: a compact
UTxO or state object whose transitions are authorised by that ledger. Lerna
need only know that the containing ledger has a deterministic escalation path
and a final UTxO-set semantics. Neither layer should depend on informal
operator promises about how many outputs will be generated at settlement time.

This also explains why a Lerna channel should not be described as a smaller
Interhead. Interhead preserves a Head-like enforceable state across two
containers and, under dispute, can move that state into the ordinary Hydra
state space. Lerna preserves rights inside one such container. Its escalation
unit is not a virtual Head conversion; it is a deterministic materialisation
batch, a receipt release, or a shutdown state that the container can include in
its next authenticated state. The difference is small in wording but large in
security accounting: Interhead reasons about Head membership, intermediary
collateral, and conversion phases; Lerna reasons about right identifiers,
touched dependencies, min-ADA reserves, and output descriptors.

\begin{invariant}[Head-like ledger containment]
A physical Head, an Interhead virtual Head, or a Hydrozoa-style ledger is a
valid container for Lerna only if it gives Lerna three explicit interfaces:
authenticated state inclusion, a bounded close or escalation path, and final
settlement as ordinary UTxOs or UTxO-equivalent withdrawals. Lerna may rely on
those interfaces, but the outer ledger is not required to interpret hidden
factory rights. Any unresolved right that could affect final ownership must
therefore be checkpointed, materialised, or receipt-released before the
container finalises.
\end{invariant}

\section{Inner rollout before outer rollout: Lerna and Hydrozoa}

Hydrozoa gives Lerna a second design reference because it separates execution
from custody~\cite{hydrozoa-spec}. In Hydrozoa's optimistic multisig regime,
the L1 script mostly sees a treasury controlled by unanimous peer signatures;
the L2 consensus protocol interprets the ledger. If peers stop co-operating,
the rule-based regime moves custody to Plutus scripts that resolve the latest
compatible L2 block and then release funds according to that state. Hydrozoa's
deterministic rollout algorithm exists because a large withdrawal set may not
fit into one Cardano transaction.

Lerna borrows the ethic, not the full mechanism. Hydrozoa rolls an L2 ledger
out to L1. Lerna rolls compact factory rights out to Head UTxOs before the
outer system settles. This is inner rollout. It happens below Hydra fanout and
below a Hydrozoa-style outer rollout:
\[
\begin{array}{c}
\text{compact Lerna rights}\\
\downarrow\\
\text{ordinary Head UTxOs}\\
\downarrow\\
\text{Hydra fanout or Hydrozoa-style L1 rollout}
\end{array}
\]

The ordering matters. If compact factory state is left unresolved until the
outer channel is already closing, peers must negotiate the shape, order,
min-ADA, datum policy, and native-asset layout of many outputs at precisely the
moment when liveness is weakest. Lerna therefore requires a deterministic
materialisation plan while the Head is still open. V0 keeps that plan entirely
Head-local. V1 would move some unresolved rights into a script-controlled claim
UTxO, but that is a future research mode and not a deployable fallback in this
paper.

The borrowed principle is custody clarity. In Hydrozoa, the optimistic ledger
may execute cheaply while custody remains in a treasury whose rule-based
fallback can be made explicit. In Lerna V0, virtual-channel updates may execute
cheaply while custody remains in the containing Head UTxO set. Inner
materialisation is the point where the compact right becomes a concrete output
that the outer ledger already knows how to settle. This is why Lerna is
Hydra-native rather than Hydrozoa-competitive: it uses the same conservative
exit ethic at a lower layer.

The analogy must stop there. Hydrozoa's rule-based regime is an outer custody
fallback: it moves control of the treasury to scripts that can resolve a
compatible L2 block and roll withdrawals out. Lerna V0 has no such outer
fallback of its own. Its compact factory datum is not a treasury and not a
standalone custody object. It is only safe while the containing ledger can still
include the materialisation transaction that turns compact rights into the
ordinary outputs that Hydra or Hydrozoa already knows how to settle.

\begin{proposition}[Inner-before-outer rollout]
For V0, a compact Lerna right may remain compact only while the containing Head
or Head-like ledger remains able to include a deterministic materialisation
transaction in a later authenticated state. Once outer settlement is imminent,
the right must be represented as an ordinary Head output, an agreed receipt
release, or a full-participant shutdown state. Otherwise the outer rollout is
being asked to settle a claim whose concrete Cardano output shape has not been
agreed.
\end{proposition}

\begin{proof}[Argument]
Hydra fanout and Hydrozoa-style rollout settle UTxO sets or withdrawals, not
application-private Merkle roots. If a compact right remains unresolved at the
outer boundary, the parties still need to decide its address, value, min-ADA,
datum, reference-script policy, and native-asset bundle. Those choices are
consensus-relevant because different choices create different ledger outputs.
V0 therefore treats inner materialisation as part of the state that must be
settled before the outer mechanism is invoked.
\end{proof}

\section{Why compact factory state matters for Hydra fanout}

Hydra fanout distributes UTxOs, not abstract application rights. This is the
central reason Lerna cannot stop at a compact Merkle root. A Head that closes
with a large or awkward final UTxO set must still satisfy Cardano transaction
size, execution-budget, native-asset, datum, and min-ADA limits. Hydra's known
limitations make this operational rather than cosmetic: too many participants
are bounded by configured and ledger limits; unburned native tokens minted
inside a Head can prevent finalisation; partial fanout can leave problematic
token-carrying outputs stuck; and an intrinsically oversized UTxO cannot be
fanned out merely by splitting the fanout transaction~\cite{hydra-known}.
The current documentation's warning that roughly dozens of Head UTxOs may
already create fanout pressure should be read as an engineering signal rather
than a hard protocol constant: the exact number depends on ledger parameters,
output shape, scripts, datums, and implementation changes, but the direction of
pressure is stable.

Compact factory state is Lerna's answer to fanout UTxO pressure. During normal
operation, a factory may represent many virtual channels through a small number
of Head-visible outputs. This reduces snapshot size, avoids unnecessary
Head-local output churn, and keeps local rebalances from becoming fanout debt
immediately. The answer is deliberately incomplete unless paired with
deterministic materialisation. Compactness postpones UTxO expansion; it does
not abolish it. Before Hydra fanout, every unresolved V0 right must either be
materialised into an ordinary Head UTxO, absorbed into a fresh agreed factory
state with no unilateral claim, or released by receipt.

This is why Lerna treats materialisation capacity as a safety assumption rather
than an optimisation. A factory that admits more compact rights than it can
unfold before the contestation deadline is unsafe even if every virtual-channel
state is correctly signed. The protocol must measure output count, datum size,
asset-bundle size, min-ADA reserve, and watcher response time as part of the
factory policy.

Compactness therefore has a precise meaning. It is not a promise that the final
Head state will be small. It is a promise that the fast path can keep temporary
rights compact while the shutdown path remains capable of producing the
ordinary Head UTxOs that Hydra requires. If that second promise is not measured
and enforced, Lerna merely hides fanout work until the worst moment.

\begin{invariant}[Fanout debt]
Every compact active right carries a fanout debt equal to the deterministic
Head-output work needed to settle it. A factory state is admissible only while
the sum of its current fanout debt fits within the deployment's shutdown and
contestation envelope, or while all affected parties have signed a cooperative
state that removes any unilateral claim. This invariant is the price Lerna pays
for avoiding Head-visible UTxO churn on the fast path.
\end{invariant}

\section{Compatibility with physical Heads, Interhead virtual Heads, and Hydrozoa-style ledgers}

Lerna is a rights layer, so its compatibility question is not whether it
replaces a channel protocol. The question is which outer ledger supplies
execution, custody, and final settlement while Lerna supplies compact rights
and inner materialisation.

\begin{longtable}{L{0.24\textwidth}L{0.34\textwidth}L{0.31\textwidth}}
\toprule
\textbf{Outer ledger} & \textbf{Lerna role} & \textbf{Required boundary}\\
\midrule
\endfirsthead
\toprule
\textbf{Outer ledger} & \textbf{Lerna role} & \textbf{Required boundary}\\
\midrule
\endhead
Physical Hydra Head & A compact factory UTxO or Head-local state object whose rights materialise before fanout & Uses Hydra close, contest, and fanout exactly as the outer dispute substrate; no V0 L1 dispute\\
Interhead virtual Head & A factory inside the virtual Head state, managing rights below the virtualised Head & Inner materialisation deadlines must fit inside Interhead conversion and outer Hydra contestation deadlines\\
Hydrozoa multisig-style ledger & A compact L2 ledger object whose withdrawals or local exits are deterministically unfolded & Peer consensus must checkpoint factory state and agree on the inner rollout plan before outer settlement\\
Hydrozoa rule-based-style ledger & A future claim-like object whose validator might adjudicate unresolved Lerna rights & V1 only; proof grammar, latest-state rule, and output rollout must be specified and benchmarked\\
Plain Cardano L1 & A possible V1 claim UTxO after a Head has fanned out unresolved factory state & Future work; validator cost, proof size, native-asset handling, and min-ADA safety are open\\
\bottomrule
\end{longtable}

The common rule is that Lerna must not smuggle new custody assumptions into the
outer ledger. In V0, custody remains with the Hydra Head or with a
Hydrozoa-style treasury. Lerna's authority is over compact factory rights and
their Head-local materialisation.

Compatibility is therefore an adapter problem, not a replacement problem. The
adapter binds four fields: the containing ledger identity, the state-inclusion
rule, the escalation deadline, and the final output semantics. If any of those
fields is informal, a Lerna deployment cannot claim V0 safety because the
materialisation debt has no precise place to be repaid.

\section{What Lerna deliberately does not take from each prior work}

Lerna's originality depends as much on refusal as on borrowing.

\begin{description}[leftmargin=2.7cm,style=nextline]
\item[From Interhead.]
Lerna does not take Interhead's goal of virtualising an entire Hydra Head across
Heads, nor its conversion machinery as a proof of Lerna safety. It takes the
lesson that virtual state must have an explicit escalation path and that
intermediary or operator resources must be visible.

\item[From Hydrozoa.]
Lerna does not take over Hydrozoa's treasury architecture, dynamic peer
membership, or rule-based fallback as V0 machinery. It takes the separation of
execution and custody, and the insistence that large settlement sets need
deterministic rollout rather than ad hoc construction under stress.

\item[From Hydra Head.]
Lerna does not change Hydra's Head lifecycle, participation-token logic, close,
contest, or fanout semantics. It takes Hydra as the outer substrate and accepts
its practical constraints, including fanout UTxO pressure and native-token
finalisation hazards. It also does not assume that partial fanout, decommit
workflows, or future Hydra engineering improvements remove the need for
Lerna's own deterministic materialisation plan.

\item[From eltoo.]
Lerna does not take Bitcoin transaction replacement assumptions, sighash
templates, or the slogan that latest state alone defines settlement. It takes
only the non-punitive intuition that newer valid state should supersede stale
evidence when the substrate can enforce that ordering.

\end{description}

\section{Formal Objects}

This section defines protocol objects at the design level. Encoding, CBOR
schema, validator budget, and Aiken or Plutus implementation details are
deliberately left for later specifications.

\begin{definition}[LernaFactoryConfig]
The factory configuration binds the factory to a Cardano network, a Head, a
factory identity, a participant registry, and a materialisation policy.
\begin{lstlisting}
LernaFactoryConfig {
  protocol_version
  network_id
  hydra_head_id
  factory_id
  factory_epoch
  head_participant_set_hash
  virtual_participant_registry_root
  ledger_profile              // V0-ADA, multi-asset-extension, V1-claim
  asset_registry_root
  reserve_policy_hash
  fee_budget_policy_hash
  materialisation_policy_hash
  signature_scheme_id
  domain_separator
}
\end{lstlisting}
\end{definition}

For the concrete V0 target, \texttt{ledger\_profile = V0-ADA}. The factory
validator or transaction rule must reject a factory input carrying non-ADA
assets unless the profile is explicitly upgraded. This keeps the first security
claim aligned with the prototype burden rather than with the full Cardano asset
surface.

\begin{definition}[FactoryState]
The factory state is the compact state of the Lerna factory. It is not a
Cardano transaction by itself. It is the object that Head-local transactions
update and that materialisation plans unfold.
\begin{lstlisting}
FactoryState {
  config_hash
  factory_state_number
  balances_root
  subchannels_root
  membership_root
  reserve_root
  release_receipt_root
  pending_materialisation_root
  settlement_descriptor_hash
  asset_registry_hash
  operator_budget_state
}
\end{lstlisting}
\end{definition}

\begin{definition}[VirtualChannelState]
The virtual-channel state is the latest-state object for one virtual channel.
In a payment-only prototype, \texttt{app\_state\_hash} may be empty and
balances are enough. In a dApp channel, it commits to application-specific
eUTxO or state-machine material.
\begin{lstlisting}
VirtualChannelState {
  virtual_channel_id
  factory_id
  participants
  virtual_state_number
  balances
  locked_reserve
  app_state_hash
  settlement_descriptor_hash
  expiry_or_liveness_profile
}
\end{lstlisting}
\end{definition}

\begin{definition}[LatestStateHeader]
The header is the signed authority object. It must be domain-separated from
ordinary Cardano transactions, Hydra snapshots, wallet messages, and earlier
protocol versions.
\begin{lstlisting}
LatestStateHeader {
  domain_separator
  protocol_version
  network_id
  hydra_head_id
  factory_id
  factory_epoch
  state_scope
  state_number
  previous_state_hash
  new_state_hash
  transition_family
  touched_root
  settlement_descriptor_hash
  materialisation_plan_hash
  validity_interval
  hydra_snapshot_hint
}
\end{lstlisting}
\end{definition}

\paragraph{Signature domain.}
The V0 signature scheme is Ed25519 over a canonical CBOR encoding of the
\texttt{LatestStateHeader}; the digest is \(\mathrm{Blake2b\mbox{-}256}\) of the
domain-tagged payload. The initial domain tags are:
\[
\begin{array}{ll}
\texttt{lerna:v0:latest-state-header} & \text{state supersession},\\
\texttt{lerna:v0:release-receipt} & \text{right release and anti-replay},\\
\texttt{lerna:v0:materialisation-plan} & \text{batch ordering and output plan},\\
\texttt{lerna:v0:watcher-delegation} & \text{optional delegation evidence}.
\end{array}
\]
The signer registry must state whether a key is a Hydra party key, a Cardano
payment/stake key used for off-chain messages, or an application key. These key
roles are not interchangeable.

\begin{definition}[AccessManifest and TouchedLeaves]
The access manifest is the envelope-first admission object. It declares what
kind of transition is being attempted before proof bytes are interpreted.
\begin{lstlisting}
AccessManifest {
  transition_family
  authorisation_policy_id
  affected_participants
  required_signers
  read_set_root
  write_set_root
  inserted_set_root
  removed_set_root
  dependency_set_root
  asset_delta_summary
  reserve_delta_summary
}

TouchedLeaves {
  balance_leaves
  subchannel_leaves
  membership_leaves
  reserve_leaves
  release_receipt_leaves
  absence_proofs
  membership_proofs
}
\end{lstlisting}
\end{definition}

\begin{definition}[NonInterferenceProof]
The proof must establish that all rights outside the transition footprint
remain unchanged, including indirect rights through shared reserve, membership,
exit paths, and sponsor or operator budgets. A Merkle proof alone is not a
non-interference proof.
\begin{lstlisting}
NonInterferenceProof {
  access_manifest_hash
  rights_dependency_graph_hash
  untouched_rights_commitment
  unchanged_dependency_paths
  conservation_witness
}
\end{lstlisting}
\end{definition}

\begin{definition}[SettlementDescriptor]
The descriptor is not an application runtime. It is a bounded rule set for
deriving Cardano outputs or Head-local outputs from Lerna rights.
\begin{lstlisting}
SettlementDescriptor {
  descriptor_version
  output_derivation_rules
  participant_destination_policies
  asset_allocation_rules
  min_ada_rules
  datum_and_reference_script_rules
  local_exit_rules
  rebalance_and_splice_rules
  fanout_preparation_rules
}
\end{lstlisting}
\end{definition}

\begin{definition}[FanoutMaterialisationPlan]
In V0, \texttt{claim\_utxo\_descriptor\_hash} must be empty because unresolved
factory state is not allowed to survive Head fanout. In V1 it may describe a
script-controlled L1 claim UTxO.
\begin{lstlisting}
FanoutMaterialisationPlan {
  plan_id
  source_factory_state_hash
  materialisation_mode
  output_set_hash
  claim_utxo_descriptor_hash
  unresolved_channel_set_root
  batch_index
  batch_count
  budget_profile_hash
  required_signatures
  watcher_deadline_profile
}
\end{lstlisting}
\end{definition}

\begin{definition}[ReleaseReceipt]
Release receipts prevent a common ambiguity in compact factories: after a
virtual right is materialised into an ordinary Head UTxO, the compact factory
must no longer carry a spendable copy of that right.
\begin{lstlisting}
ReleaseReceipt {
  receipt_id
  receipt_domain_separator
  released_right_id
  factory_id
  factory_epoch
  virtual_channel_id
  state_number
  right_commitment_hash
  materialisation_plan_hash
  batch_index
  output_index
  replacement_output_hash
  consumed_in_factory_state_hash
  signer_set_hash
  signature
}
\end{lstlisting}
\end{definition}

\begin{definition}[Release receipt anti-replay rule]
Let
\[
\begin{aligned}
\mathsf{receiptId}=\hash(&\texttt{lerna:v0:release-receipt},
\mathsf{networkId},\mathsf{hydraHeadId},\\
&\mathsf{factoryId},\mathsf{factoryEpoch},\mathsf{releasedRightId},
\mathsf{stateNumber},\\
&\mathsf{materialisationPlanHash},\mathsf{batchIndex},
\mathsf{outputIndex}).
\end{aligned}
\]
A receipt is valid only if this identifier is absent from the old
\(\mathsf{release\_receipt\_root}\), present in the new
\(\mathsf{release\_receipt\_root}\), the released right is absent from the new
active-right commitments, and the referenced materialised output hashes exactly
to \(\mathsf{replacementOutputHash}\). The receipt signature uses the
\texttt{lerna:v0:release-receipt} domain and the same signer policy that
authorised the materialisation. A receipt cannot be replayed across Heads,
factory epochs, batch positions, or replacement outputs.
\end{definition}

\section{Head-local Validator Sketch}

The first implementation target should be an ADA-only Head-local factory
validator, or a validator-compatible transaction rule used by a Hydra
application. This does not modify Hydra protocol internals. The following is
intentionally Aiken-like pseudocode rather than production Aiken~\cite{aiken}.
Its purpose is to show the V0 prototype direction and the minimum checks that
belong on the compact factory UTxO or equivalent Head-local rule.

\begin{lstlisting}
validator lerna_factory {
  spend(datum: FactoryDatum, redeemer: FactoryRedeemer, ctx: ScriptContext) {
    let old = datum.state
    let new = redeemer.new_state
    let manifest = redeemer.access_manifest
    let touched = redeemer.touched_leaves
    let receipts = redeemer.release_receipts

    check_factory_identity(old.config_hash, new.config_hash)
    check_head_binding(old.hydra_head_id, ctx)
    check_state_number_increases(old.state_number, new.state_number)
    check_ledger_profile_is_ada_only(old, new, ctx)

    check_domain_separated_signatures(
      redeemer.latest_state_header,
      manifest.required_signers,
      redeemer.signatures,
    )

    check_manifest_commits_to_body(manifest, redeemer.proof_body)
    check_touched_roots_recompute(old, new, manifest, touched)

    if manifest.required_signers != all_factory_participants {
      reject_unless_non_interference_profile_is_active()
      check_non_interference(manifest, touched, redeemer.non_interference_proof)
    }

    check_release_receipts_are_fresh_and_consuming(old, new, receipts)
    check_ada_conservation(old, new, redeemer.materialised_outputs)
    check_min_ada_for_materialised_outputs(redeemer.materialised_outputs)
    check_operator_budget_boundary(old, new, manifest)
  }
}
\end{lstlisting}

This sketch is deliberately narrow. It is not a Hydra protocol modification and
does not claim that the V1 L1 claim validator is feasible. It only identifies
the V0-ADA checks that a prototype should measure inside a Hydra-local
transaction: identity, monotonicity, signature domain, access-manifest binding,
root recomputation, full-authorisation fallback, release-receipt consumption,
ADA conservation, min-ADA constraints, and operator-budget separation. Native
asset conservation remains an extension check, not part of the first V0 claim.

\section{Minimal ADA-only V0 State Machine}

The V0 prototype should be specified as a small state machine before any
multi-asset or V1 claim machinery is attempted. A factory state carries:
\[
F=(c,n,m,B,R,Q,E,\rho,\pi,\omega),
\]
where \(c\) is the configuration hash, \(n\) the factory state number, \(m\) the
mode, \(B\) the balance/right commitments, \(R\) the reserve commitments, \(Q\)
the pending materialisation commitment, \(E\) the release-receipt commitment,
\(\rho\) the participant registry commitment, \(\pi\) the settlement descriptor
hash, and \(\omega\) the operator-budget state. In V0-ADA the value lanes are
lovelace only, split into business balances, min-ADA reserve, shared reserve,
and operator budget.

\[
m \in
\{\mathsf{Init},\mathsf{Open},\mathsf{Checkpointing},\mathsf{Materialising},
\mathsf{ShutdownPreparing},\mathsf{Settled},\mathsf{Aborted}\}.
\]

\begin{longtable}{L{0.22\textwidth}L{0.25\textwidth}L{0.40\textwidth}}
\toprule
\textbf{Transition} & \textbf{Mode change} & \textbf{Admissibility summary}\\
\midrule
\endfirsthead
\toprule
\textbf{Transition} & \textbf{Mode change} & \textbf{Admissibility summary}\\
\midrule
\endhead
\(\mathsf{initFactory}\) & \(\mathsf{Init}\rightarrow\mathsf{Open}\) &
Head id, factory id, descriptor, signer registry, and ADA-only ledger profile
are bound; initial ADA conservation holds.\\
\(\mathsf{openVC}\) & \(\mathsf{Open}\rightarrow\mathsf{Open}\) &
Required signers authorise a new virtual right; reserve and min-ADA accounting
increase consistently; no native asset is admitted in V0-ADA.\\
\(\mathsf{checkpoint}\) & \(\mathsf{Open}\rightarrow\mathsf{Checkpointing}
\rightarrow\mathsf{Open}\) & A domain-separated latest-state header strictly
increases the relevant state number and is included in a Hydra snapshot.\\
\(\mathsf{rebalance}\) & \(\mathsf{Open}\rightarrow\mathsf{Open}\) &
The access manifest declares both source and destination rights; ADA is
conserved across affected rights and reserve lanes.\\
\(\mathsf{localExit}\) & \(\mathsf{Open}\rightarrow\mathsf{Materialising}\) &
Latest evidence, touched leaves, and either full factory authorisation or an
active non-interference proof authorise materialisation of selected rights.\\
\(\mathsf{materialiseBatch}\) &
\(\mathsf{Materialising}\rightarrow\mathsf{Materialising}\) or
\(\mathsf{Open}\) & The canonical batch creates ordinary Head outputs, writes
fresh release receipts, removes active compact rights, and updates the pending
suffix commitment.\\
\(\mathsf{prepareShutdown}\) &
\(\mathsf{Open}\rightarrow\mathsf{ShutdownPreparing}\) & Materialisation debt
exceeds the configured threshold or the Head is expected to close; new compact
opens stop.\\
\(\mathsf{cooperativeSettle}\) &
Open or shutdown-preparing to settled &
All remaining rights are materialised, released, or covered by full-participant
shutdown signatures; no unresolved unilateral claim remains.\\
\(\mathsf{abortV0}\) & any non-settled mode \(\rightarrow\mathsf{Aborted}\) &
Entered only as an analytical state when the liveness bridge or materialisation
capacity assumption is violated; V0 specifies no independent recovery from it.\\
\bottomrule
\end{longtable}

\begin{definition}[Transition validity]
\(\mathsf{ValidV0ADA}(F,F',T,W)\) holds when the transition \(T\) and witness
\(W\) satisfy factory identity, Head binding, strict state-number monotonicity,
signature-domain correctness, manifest-to-body binding, touched-root
recomputation, release-receipt freshness, ADA conservation, min-ADA reserve
sufficiency for every materialised output, and the operator-budget boundary.
If \(T\) uses fewer than all factory signers, validity additionally requires an
active and benchmarked \(\mathsf{ValidNonInterference}\) predicate; otherwise
the transition is invalid.
\end{definition}

\begin{remark}[Same-number equivocation]
For a fixed \((\mathsf{network},\mathsf{head},\mathsf{factory},
\mathsf{epoch},\mathsf{scope})\), two different headers with the same
\(\mathsf{stateNumber}\) and different \(\mathsf{newStateHash}\) are an
equivocation event. V0 does not try to pick a winner. The affected scope must
freeze until the parties sign a strictly higher repair state or the factory
falls back to full-participant shutdown. This avoids silently accepting
``highest number wins'' when the number is tied.
\end{remark}

\section{Protocol Lifecycle}

\subsection{Open Head}

Participants open a Hydra Head under the ordinary Hydra lifecycle. In the
current practical Hydra workflow, Lerna should treat Head opening, deposit, and
decommit as outer protocol operations rather than as Lerna-specific actions.
Lerna does not modify this process.

\subsection{Initialise Lerna Factory}

The Head participants create an initial Lerna factory state inside the open
Head. A conservative prototype may represent it as one script-controlled Head
UTxO with an inline datum containing the \texttt{FactoryState} hash and enough
data to update it. More compact implementations may store only the roots and
keep proofs off-chain.

Initialisation must bind the Cardano network id, Hydra Head id, Lerna factory
id, participant registry root, asset registry root, settlement descriptor hash,
and materialisation policy.

\subsection{Open Virtual Channel}

Opening a virtual channel updates the factory state. The transition touches
membership, reserve, balances, and subchannel roots. It requires signatures
from the virtual-channel participants and any factory participants whose
reserve or exit rights are affected. If the opening changes shared reserve or
membership in a way that affects everyone, full factory authorisation is
required unless non-interference proves otherwise.

\subsection{Off-chain or Head-local Update}

There are two update styles. In the first, virtual-channel participants exchange
signed latest-state headers off-chain and only submit to the Head
when they need to checkpoint, rebalance, exit, or respond to conflict. In the
second, a compact factory UTxO is updated by a transaction inside the Head, and
the resulting Head snapshot records the new factory roots.

The V0 prototype should support Head-local checkpointing first, then add purely
off-chain virtual updates once materialisation and release receipts are well
tested.

\subsection{Local Rebalance, Splice, and Exit}

A rebalance moves value between virtual channels without changing the outer
Head deposit. A splice changes the compact factory reserve by depositing into,
decommitting from, or rearranging Head-local value. Lerna must distinguish
these carefully: rebalance changes rights inside the factory; splice changes
the factory's Head-visible value or materialisation reserve; neither may
silently change the factory id or replay domain.

A local exit lets a participant materialise a virtual-channel right without
closing the entire factory. It requires latest-state evidence for the relevant
virtual channel, release receipts for rights removed from compact state,
touched leaves and membership proofs, non-interference proof for untouched
participants or full factory authorisation, and exact output materialisation
under the settlement descriptor.

\subsection{Materialise Inside Head}

Materialisation converts compact factory rights into ordinary Head UTxOs. In
V0 this is the main safety operation. Before Head close, every unresolved
virtual channel must either materialise into ordinary Head outputs, be
reabsorbed into the factory with fresh signatures, be closed with release
receipts, or be covered by a full-participant shutdown plan.

The materialised outputs must satisfy Cardano eUTxO constraints, including
min-ADA and native-asset accounting.

\subsection{Head Close, Contest, and Fanout}

If the Head is closed, Hydra's own close and contest rules decide which Head
snapshot is final. Lerna V0 does not create a separate L1 dispute. Therefore,
Lerna participants must ensure the final Head snapshot contains either ordinary
materialised outputs or a fully agreed factory shutdown state that has no
unresolved unilateral claims.

If a stale Head snapshot omits a newer Lerna materialisation, the answer is not
a Lerna-specific L1 action. The answer is to contest the Head with a newer
Hydra snapshot that includes the correct Lerna state or materialised outputs.
This is why watcher responsibilities must cover both Hydra snapshots and Lerna
latest-state evidence.

\subsection{V1 Claim-UTxO Fanout}

V1 is future work. If the Head closes before all virtual channels are
materialised, the final Head UTxO set may include a script-controlled L1 claim
UTxO representing the unresolved factory state. An Aiken or Plutus validator
would then enforce domain-separated latest-state signatures, state
monotonicity, local exits, non-interference proofs, native-asset conservation,
min-ADA-safe output creation, and release receipt consumption.

This is research-heavy. Until such validators are specified and tested, V1
should not be presented as deployable.

\section{Latest-state Settlement Semantics}

Lerna uses monotonic state numbers. For a fixed
\[
(\mathsf{network},\mathsf{head},\mathsf{factory},\mathsf{epoch},\mathsf{scope}),
\]
a valid state with number \(n+k\), \(k>0\), supersedes a valid state with number
\(n\) at the Lerna layer if the header is domain-separated, signatures match
the required signing set, the transition family is admissible, touched leaves
recompute the old and new roots, conservation checks pass, and release receipts
prevent double materialisation.

There is no penalty mechanism. A participant presenting stale evidence does not
lose funds by punishment. The safety objective is simply that stale evidence
cannot finalise if newer valid evidence is available through the appropriate
Hydra contest or Lerna materialisation path.

\begin{lemma}[Hydra inclusion lemma for V0]
Assume an honest authorised actor observes a newer valid Lerna state \(S'\)
before the Hydra contestation deadline, can construct a Head transaction that
checkpoints or materialises \(S'\), obtains a Hydra snapshot containing that
transaction, and successfully contests a stale close with that newer snapshot.
Then the final Hydra fanout follows the newer snapshot rather than the stale
snapshot.
\end{lemma}

\begin{proof}[Argument]
This is a restatement of the outer Hydra close/contest mechanism applied to a
Head state that happens to contain Lerna data. In V0, Lerna has no independent
L1 dispute mechanism. Once \(S'\) is included in a newer valid Hydra snapshot,
Hydra's contest rule is the mechanism that prevents the older snapshot from
being final. The lemma proves only the inclusion bridge; it does not prove that
a non-Head virtual participant can force inclusion, that proof material will be
available, or that materialisation will fit in the remaining window.
\end{proof}

\begin{remark}[Not a Lerna security theorem]
The lemma should not be cited as a standalone Lerna latest-state safety theorem.
The hard Lerna obligations are exactly the premises: authorised inclusion,
proof availability, materialisation capacity, receipt consumption, and timely
contest. A deployable claim would need to make those premises mechanisms rather
than assumptions.
\end{remark}

\section{Factory-local Proof Mode}

A Lerna factory may contain many balances and subchannels. Requiring all
participants to sign every local action is safe but undermines the purpose of a
factory. Requiring only affected participants to sign is unsafe unless the
protocol proves that untouched participants are not affected.

The proof mode has four layers:

\begin{enumerate}[label=(\arabic*),leftmargin=2.2em]
\item commitment roots for balances, subchannels, membership, reserve, and
release receipts;
\item an access manifest declaring the transition family and footprint before
proof interpretation;
\item touched leaves for the items read, written, inserted, removed, or proven
absent;
\item a non-interference proof showing that untouched rights remain unchanged,
including dependency paths through shared reserve and membership.
\end{enumerate}

\begin{definition}[Rights dependency graph]
For a factory state \(F\), the rights dependency graph is a labelled directed
graph
\[
G_F=(V,E,\lambda)
\]
whose vertices are right-bearing atoms:
\[
V =
V_{\mathsf{balance}} \cup V_{\mathsf{reserve}} \cup V_{\mathsf{membership}}
\cup V_{\mathsf{descriptor}} \cup V_{\mathsf{receipt}} \cup
V_{\mathsf{operator}}.
\]
An edge \(u\rightarrow v\) means that changing \(u\) may change the value,
exit path, materialisation viability, or authorisation condition of \(v\). For
example, a shared reserve vertex points to every balance right that depends on
that reserve; a membership vertex points to the signing policy and local-exit
rights that use that member set; a descriptor vertex points to every right
whose materialised output it derives. The graph commitment is part of the
factory state, so a transition cannot choose its dependency graph after seeing
which proof is convenient.
\end{definition}

\begin{definition}[Touched closure]
For a transition with access manifest \(M\), let \(A(M)\) be the set of vertices
declared read, written, inserted, removed, or proven absent. The touched
closure is
\[
\mathsf{cl}_{G_F}(M)=
A(M)\cup\{v \mid \exists u\in A(M)\ \text{and a path }u\leadsto v
\text{ in }G_F\}.
\]
Reduced authorisation is admissible only for participants whose rights are
inside this closure. Rights outside the closure must have unchanged membership,
balance, reserve, descriptor, and exit-path commitments.
\end{definition}

\begin{definition}[Valid non-interference predicate]
\(\mathsf{ValidNonInterference}(F,F',M,W)\) holds only if \(W\) proves:
\begin{enumerate}[label=(\alph*),leftmargin=2.2em]
\item the committed dependency graph used for \(F\) and \(F'\) is the graph
declared by \(M\);
\item all writes, insertions, removals, and absence claims are contained in
\(\mathsf{cl}_{G_F}(M)\);
\item every vertex outside \(\mathsf{cl}_{G_F}(M)\) has the same committed
right in \(F\) and \(F'\);
\item ADA conservation holds over the closure plus any authorised splice,
operator-budget, or materialised-output lanes;
\item the reduced signer set covers every participant whose right lies inside
the closure or whose authorisation policy is changed by the transition.
\end{enumerate}
If any of these checks cannot be represented by the chosen proof system, V0
must reject the reduced-signature transition and require full factory
authorisation.
\end{definition}

\begin{invariant}[Envelope-first admissibility]
A proof body is interpreted only after the access manifest has fixed the
transition family and committed to that body. The proof bytes do not get to
choose their own authority class.
\end{invariant}

\begin{invariant}[Locality is a proof obligation]
If non-interference cannot be established for a reduced-signature transition,
the transition falls back to full factory authorisation.
\end{invariant}

\begin{figure}[htbp]
\centering
\begin{tikzpicture}[
  font=\footnotesize,
  box/.style={draw, rounded corners=2pt, align=center, text width=2.7cm,
              minimum height=0.65cm, inner sep=4pt},
  wide/.style={draw, rounded corners=2pt, align=center, text width=3.1cm,
               minimum height=0.78cm, inner sep=5pt},
  arrow/.style={-Latex, thick, draw=LernaGrey}
]
\node[box, fill=LernaBlue!7, draw=LernaBlue] (balances) {balances root};
\node[box, fill=LernaBlue!7, draw=LernaBlue, below=0.35cm of balances] (subs) {subchannels root};
\node[box, fill=LernaBlue!7, draw=LernaBlue, below=0.35cm of subs] (members) {membership root};
\node[box, fill=LernaBlue!7, draw=LernaBlue, below=0.35cm of members] (reserve) {reserve root};

\node[wide, fill=LernaAmber!12, draw=LernaAmber, right=1.15cm of subs] (manifest) {access manifest\\transition family};
\node[wide, fill=LernaAmber!8, draw=LernaAmber, right=1.15cm of reserve] (proof) {non-interference\\dependency proof};
\node[wide, fill=LernaGreen!8, draw=LernaGreen, right=1.25cm of manifest] (exit) {local exit\\or update};
\node[wide, fill=gray!8, draw=gray!70, right=1.25cm of proof] (signers) {required\\signatures};

\draw[arrow] (balances.east) -- (manifest.west);
\draw[arrow] (subs.east) -- (manifest.west);
\draw[arrow] (members.east) -- (proof.west);
\draw[arrow] (reserve.east) -- (proof.west);
\draw[arrow] (manifest.east) -- (exit.west);
\draw[arrow] (proof.east) -- (signers.west);
\draw[arrow] (signers.north) -- (exit.south);
\end{tikzpicture}
\caption{Factory proof mode. The touched footprint and the non-interference
argument are different objects. A balance proof is not enough when reserve,
membership, or exit rights depend on that balance.}
\end{figure}

\section{Conservation Model}

The first V0 ledger profile is ADA-only. Native assets are described here only
as an extension target. This downgrade matters: Cardano native assets are
powerful, but Hydra's known limitations around unburned tokens and large output
bundles make native-asset fanout an unsafe first claim.

\subsection{ADA-only V0 Equations}

For a factory state \(F\), define:
\[
\begin{array}{ll}
B(F) & \text{sum of spendable lovelace owed by active virtual rights},\\
M(F) & \text{sum of lovelace reserved only for min-ADA of materialised outputs},\\
R(F) & \text{shared factory reserve not assigned to a single virtual balance},\\
O(F) & \text{operator budget authorised for fees and publication costs},\\
P(F) & \text{pending materialised-output lovelace committed by the current plan}.
\end{array}
\]

For an ordinary internal update with no Head-visible splice and no
materialisation,
\[
B(F)+M(F)+R(F)+O(F)+P(F)
=
B(F')+M(F')+R(F')+O(F')+P(F').
\]

For a materialisation batch producing ordinary Head outputs \(Y\), let
\(\mathrm{ada}(Y)\) be the lovelace in those outputs and let
\(\mathrm{fee}_O\) be the operator-budget amount explicitly consumed by the
transition. The admissibility equation is:
\[
B(F)+M(F)+R(F)+O(F)+P(F)
=
B(F')+M(F')+R(F')+O(F')+P(F')+\mathrm{ada}(Y)+\mathrm{fee}_O.
\]
Every output \(y\in Y\) must satisfy the settlement descriptor and its min-ADA
requirement; the min-ADA lane is not spendable business balance. A local exit
that pays a participant by stealing from \(M(F)\), \(R(F)\), or \(O(F)\) is
invalid unless the transition family and signer set explicitly authorise the
corresponding lane movement.

\begin{invariant}[ADA-only conservation]
For every admitted V0-ADA transition, the equations above hold, every created
release receipt consumes exactly one active right, and the same right is not
present in both \(B(F')\) and a materialised Head output.
\end{invariant}

\subsection{Native-asset Extension}

Cardano native assets are identified by policy ID and asset name. A future
multi-asset Lerna profile must check a separate conservation equation for each
asset lane
\[
a=(\mathsf{policyId},\mathsf{assetName}),
\qquad
Q_a(F)+I_a-M_a-Burn_a = Q_a(F')+Y_a,
\]
where \(Q_a\) is the quantity committed in compact active rights and reserves,
\(I_a\) is an explicitly authorised Head-local deposit or splice input,
\(M_a\) and \(Burn_a\) are authorised mint/burn quantities under the relevant
policy, and \(Y_a\) is the quantity emitted into materialised outputs. Metadata,
ticker symbols, and display names are not asset identity. This equation is not
part of the ADA-only V0 prototype claim.

The following table contrasts ordinary accounting intuition with Lerna's
required treatment.

\begin{longtable}{L{0.24\textwidth}L{0.31\textwidth}L{0.34\textwidth}}
\toprule
\textbf{Lane} & \textbf{What Lerna tracks} & \textbf{Main hazard}\\
\midrule
\endfirsthead
\toprule
\textbf{Lane} & \textbf{What Lerna tracks} & \textbf{Main hazard}\\
\midrule
\endhead
Lovelace balance & Spendable ADA owed to virtual participants & Accidentally using business balance to pay operator costs\\
Min-ADA reserve & Lovelace required to make materialised outputs valid & Double-counting output viability reserve as spendable balance\\
Native asset & Policy ID, asset name, and quantity & Treating ticker, metadata, or display name as asset identity\\
Factory reserve & Shared liquidity and exit reserve claims & Local transition affecting untouched participants through shared reserve\\
Operator budget & Fees and operational close/materialisation costs & Hidden leakage from virtual-channel balances\\
Release receipts & Rights already materialised or closed & Double claim through compact and materialised forms\\
\bottomrule
\end{longtable}

Lerna's partitioned-conservation rule is a native Cardano accounting discipline:
Cardano fees, min-ADA, native-asset bundles, operator budgets, and participant
balances must be tracked separately. Publication or operation costs must not
invisibly leak from participant balances.

\section{Deterministic Materialisation}

Hydrozoa's deterministic rollout algorithm is a useful warning: a state-channel
paper that does not specify how large withdrawal sets become Cardano
transactions has not finished its exit design. Lerna adopts the same design
ethic but applies it one layer inward. The question is not only how a Head fans
out; it is how compact factory rights become a Head UTxO set before that
fanout.

Materialisation must be measurable. A V0 prototype should report, for a chosen
Hydra contestation period, the number of rights unfolded, output count, datum
size, min-ADA footprint, native-asset bundle size, transaction count, witness
size estimate, snapshot-confirmation latency, and observed
time-to-materialise. Without those measurements, compactness is only a
presentation choice, not a safety argument.

This measurement is not a benchmark afterthought. It is part of the admission
rule for opening new compact rights. A deployment that cannot estimate the
remaining materialisation workload should refuse new virtual-channel opens and
enter shutdown preparation before the Head is under close pressure. In
particular, a factory that admits more compact rights than it can unfold within
its chosen shutdown envelope is unsafe even when all relevant signatures are
valid.

\begin{definition}[Materialisation equivalence]
A materialised Head output is equivalent to a compact Lerna right if its
address, value, datum policy, reference-script policy, and release receipt match
the settlement descriptor for that right. Equal value alone is insufficient.
\end{definition}

\subsection{V0 Algorithm}

Given a source factory state \(F\), a settlement descriptor \(D\), and a set of
rights \(R\) to unfold, V0 materialisation proceeds under a committed
\(\mathsf{BudgetProfile}\). The profile contains conservative constants:
\[
\mathsf{BudgetProfile}=(B_{\mathsf{bytes}},B_{\mathsf{ex}},
B_{\mathsf{outputs}},B_{\mathsf{datum}},B_{\mathsf{value}},B_{\mathsf{margin}},
S_{\mathsf{witness}}),
\]
where the final component is a fixed witness-size estimate for the selected
signing policy. This profile is part of the factory configuration or the
materialisation policy hash, so peers compute the same batches offline.

\begin{enumerate}[label=(\arabic*),leftmargin=2.2em]
\item Sort \(R\) by canonical right identifier:
\[
(\mathsf{factoryId}, \mathsf{virtualChannelId}, \mathsf{rightKind},
\mathsf{participantId}, \mathsf{nonce}).
\]
\item For each right, derive a candidate Head output using \(D\). The derived
output includes address, lovelace, datum, optional reference-script
expectation, and min-ADA reserve. Under \(\mathsf{V0\mbox{-}ADA}\), reject the
plan if any candidate output contains non-ADA assets.
\item Reject the plan if any output fails Cardano output validity, min-ADA, or
value-bundle constraints.
\item Attach one release receipt per materialised right and update the
\(\mathsf{release\_receipt\_root}\).
\item Define
\[
\mathsf{Fits}_P(k)=\mathsf{true}
\]
exactly when the first \(k\) candidate outputs, receipts, datum updates, and
fixed witness estimate fit within the budget profile \(P\); otherwise
\(\mathsf{Fits}_P(k)=\mathsf{false}\).
The batch size is
\[
k^\star=\max\{k\mid 0<k\leq |R|,\ \mathsf{Fits}_P(k)\}.
\]
If no such \(k\) exists, the plan is invalid and the factory must enter
shutdown-preparation or full-authorisation recovery.
\item Materialise exactly the first \(k^\star\) candidate outputs. The new
factory state commits to the untouched suffix \(R[k^\star+1..]\) in
\(\mathsf{pending\_materialisation\_root}\), increments
\(\mathsf{batchIndex}\), and records the fresh release receipts for the prefix.
\item Repeat deterministically until the suffix is empty or the shutdown
deadline makes further cooperative progress unsafe.
\end{enumerate}

\paragraph{Batch failure semantics.}
Only a confirmed Head-local transaction changes the factory state. If an
off-chain preflight, signature collection, or transaction submission for batch
\(i\) fails, the last confirmed factory state remains authoritative and the
same batch can be reconstructed from \((F,D,R,\mathsf{BudgetProfile},i)\). If
batches \(0..i-1\) have already been confirmed and the Head closes before batch
\(i\) is confirmed, the pending suffix remains compact. In V0 that suffix has
no independent L1 claim; the deployment is safe only if a newer Hydra snapshot
including further materialisation can still be contested, or if all remaining
rights are covered by full-participant shutdown signatures.

\paragraph{Concurrency rule.}
A factory may have at most one active materialisation plan for a given
\((\mathsf{factoryId},\mathsf{factoryEpoch},\mathsf{sourceFactoryStateHash})\).
The \(\mathsf{planId}\), \(\mathsf{batchIndex}\), and release-receipt ids make
concurrent duplicate plans conflict rather than silently create two claims for
the same right.

This is not identical to Hydrozoa rollout. Hydrozoa rolls L2 withdrawals out to
L1 through settlement/finalisation and rollout UTxOs. Lerna V0 rolls compact
factory rights out to ordinary Head UTxOs before Hydra fanout. The shared idea
is determinism: peers should not have to negotiate transaction minutiae at the
moment of exit.

\subsection{Shutdown Preparation}

A Lerna deployment should define a shutdown-preparation threshold
\(\Theta_{\mathsf{mat}}\). If the expected materialisation work cannot be
completed with margin before the Hydra contestation deadline, the factory must
stop accepting new compact virtual-channel opens and begin unfolding rights.
The threshold depends on:

\[
\Theta_{\mathsf{mat}} =
f(N_{\mathsf{rights}}, S_{\mathsf{datum}}, S_{\mathsf{assets}},
T_{\mathsf{head}}, T_{\mathsf{watch}}, B_{\mathsf{tx}}),
\]

where \(N_{\mathsf{rights}}\) is the number of rights to unfold,
\(S_{\mathsf{datum}}\) is the expected datum footprint,
\(S_{\mathsf{assets}}\) is the native-asset bundle footprint,
\(T_{\mathsf{head}}\) is the Head contestation period,
\(T_{\mathsf{watch}}\) is watcher response time, and \(B_{\mathsf{tx}}\) is the
transaction budget profile. This paper does not know \(f\) in advance and does
not pretend that one symbolic function settles the engineering problem. It
requires each deployment to measure a conservative bound for its own ledger
profile. The safety question is whether the current compact state can be
unfolded with margin, not whether its signatures are well formed in isolation.

\subsection{V1 Claim-UTxO Discipline}

V1 does not remove the need for deterministic materialisation. It moves part of
the process to L1. A claim UTxO must still define a deterministic withdrawal
order, a proof grammar, a state-number rule, and a way to split large output
sets across multiple transactions. Otherwise V1 merely postpones the exit
problem until the worst possible moment.

\section{Hydra L2/L3 Interaction}

Lerna is naturally an inner layer. At L1, Cardano provides settlement,
validators, native assets, and finality. At L2, Hydra or a Hydrozoa-style Head
provides fast shared execution and snapshot/fanout safety. Inside the Head,
Lerna provides virtual-channel factory semantics. Above Lerna, applications can
run payment channels, app-specific channels, local rebalances, and conditional
exits.

This can be described as L3 only if the term is used carefully. Lerna is not a
new global consensus layer above Hydra. It is a factory discipline whose
ordinary state transitions can be off-chain relative to the Head and whose
checkpoints can be Head-local.

Compatibility with Interhead and virtual Head directions is plausible but not
solved. A virtual channel could connect liquidity positions represented in two
Heads, but then Lerna needs cross-Head evidence, timeout alignment, and
liquidity-discovery logic. Those are out of scope.

Hydrozoa dynamic membership may make Lerna more useful, not less. If Head
membership can evolve, the factory participant registry and rights-dependency
graph become even more important. Lerna should not assume static Head
membership forever, but V0 should begin with a fixed participant set.

\section{Security Properties}

This section states target safety properties and the mechanism expected to
enforce each one. The term ``safety property'' is used only where an enforcement
mechanism is stated.

\begin{description}[leftmargin=2.8cm,style=nextline]
\item[Anti-replay.]
State signatures cannot be replayed across networks, Heads, factories, factory
epochs, state scopes, or descriptor versions because \texttt{LatestStateHeader}
binds those fields under a domain separator. Assumptions: signature
unforgeability, canonical encoding, and no domain-separator collision.

\item[State monotonicity.]
For a given state scope, valid transitions strictly increase
\texttt{state\_number}. V0 enforces this through Head-local factory update rules
and by requiring any contested Head snapshot to include the latest valid Lerna
checkpoint or materialisation.

\item[Virtual-channel balance conservation.]
In V0-ADA, lovelace balances are conserved per virtual channel unless a
transition explicitly moves value through a declared rebalance, splice, deposit,
exit, or release. Mechanism: access manifest, touched leaves, reserve delta
checks, settlement descriptor, and the ADA equations in the conservation model.

\item[Factory non-interference.]
Untouched participants' rights remain unchanged when a reduced signing set
authorises a local action. Mechanism: rights-dependency graph,
non-interference proof, and fallback to full authorisation when proof is
insufficient.

\item[Safe materialisation.]
Compact rights unfold into valid Cardano outputs exactly once. Mechanism:
settlement descriptor, min-ADA rules, materialisation plan, and release
receipts.

\item[Close/fanout safety.]
In V0, no unresolved unilateral virtual claim should survive Head fanout.
Mechanism: materialisation-before-fanout rule and Hydra contestation with newer
snapshots when stale Head snapshots omit required Lerna state.

\item[No accidental Head-level UTxO explosion.]
The factory normally remains compact inside the Head. Mechanism: compact
\texttt{FactoryState} plus selective materialisation. The protocol deliberately
postpones UTxO expansion until local exit, rebalance boundary, or Head shutdown
preparation.

\item[Native-asset extension no-drift target.]
In a future multi-asset profile, native assets must not appear or disappear
merely because a compact proof changes a root. Proposed mechanism: per-policy
asset conservation, explicit mint/burn policy checks when relevant, and
materialised-output equality under the descriptor. This is an extension target,
not part of the ADA-only V0 claim.
\end{description}

\section{Failure Scenarios}

The following scenarios are part of the V0 specification, not edge cases to be
hidden in implementation notes.

\begin{longtable}{L{0.25\textwidth}L{0.34\textwidth}L{0.30\textwidth}}
\toprule
\textbf{Scenario} & \textbf{What happens in V0} & \textbf{Design consequence}\\
\midrule
\endfirsthead
\toprule
\textbf{Scenario} & \textbf{What happens in V0} & \textbf{Design consequence}\\
\midrule
\endhead
Operator refuses to checkpoint & A virtual participant who is not a Head
participant cannot force inclusion through Lerna alone & Deployment must name
an authorised fallback actor or admit assisted, not unilateral, security\\
Head participant censors materialisation & If no honest authorised participant
can create a newer snapshot, stale Head state may finalise & V0 fails; V1
claim-UTxO or different custody model is required\\
Watcher is late & The newer Lerna state may be valid but irrelevant to final
fanout if the contest deadline passes & Watcher response time is part of the
materialisation threshold, not an operational afterthought\\
Proof material withheld & A signed high-number header without touched leaves,
descriptor body, or receipt witnesses cannot be validated & Header signing
should bind proof availability policy; otherwise freeze and require repair or
full authorisation\\
Same-number equivocation & Two different headers at the same scope and number
do not have an ordering relation & Freeze the affected scope until a strictly
higher repair state or full shutdown is signed\\
Too many outputs & Deterministic materialisation cannot complete with margin &
Stop accepting compact opens and enter shutdown preparation before close risk\\
Receipt replay attempt & A right is presented both as compact state and as a
materialised output & Reject because the receipt id must be fresh in the old
state, present in the new state, and tied to the replacement output\\
Native-asset bundle appears in V0-ADA & The factory input or materialised output
contains a non-ADA asset & Reject under \texttt{ledger\_profile = V0-ADA};
multi-asset requires a separate profile and benchmark\\
\bottomrule
\end{longtable}

\section{Limitations and Open Questions}

Hydra fanout constraints are central. A Lerna plan that materialises too many
outputs too late may fail in practice even if it is theoretically valid.
Partial fanout helps but does not remove intrinsic Cardano transaction-size
limits, especially for large datums or awkward native-asset bundles.

Head participant limits also matter. Lerna can reduce virtual-channel signing
sets, but the outer Head still has a participant set and networking topology.
Non-Head users need representation and watcher arrangements.

V1 claim UTxOs are complex. A validator must handle signatures, monotonicity,
local exits, non-interference, asset conservation, min-ADA, and proof sizes
within Cardano execution budgets. This paper does not claim that such a
validator is complete.

Proof size is unknown. Sparse Merkle proofs, KZG-style accumulators, or other
commitments may be useful, but the choice must be benchmarked against Aiken or
Plutus costs and Head-local transaction ergonomics.

Routing and liquidity discovery are out of scope. Lerna defines factory-local
safety, not a payment network. Privacy is mostly out of scope. Commitment roots
avoid full plaintext disclosure in factory mode, but access manifests and
touched leaves still leak information. Production readiness is not claimed. The
immediate output should be a prototype and a test agenda.

\section{Evaluation Agenda}

The first evaluation should be small and concrete.

\subsection{Minimal Hydra Local Demo}

Run a local Hydra Head with two or three participants. Create a compact Lerna
factory UTxO inside the Head. Execute virtual-channel open, update, local exit,
and materialisation transactions. Close and fan out only after materialisation.

\subsection{Minimum Prototype Measurements}

The first report should include the following table for the ADA-only path. This
paper does not invent measurements; it defines the minimum data needed before a
stronger claim is credible.

\begin{longtable}{L{0.30\textwidth}L{0.28\textwidth}L{0.31\textwidth}}
\toprule
\textbf{Measurement} & \textbf{Prototype case} & \textbf{Why it matters}\\
\midrule
\endfirsthead
\toprule
\textbf{Measurement} & \textbf{Prototype case} & \textbf{Why it matters}\\
\midrule
\endhead
Head-local transaction bytes & open, checkpoint, local exit, materialise batch
& Determines the safe batch budget\\
Validator execution units & compact factory spend & Shows whether Aiken/Plutus
checks are plausible\\
Materialised output count per batch & ADA-only exits & Bounds shutdown
preparation\\
Inline datum bytes & compact factory datum and materialised outputs & Detects
unfannable large-output risk\\
Min-ADA locked for materialisation & two-channel local exit & Separates
business balance from output viability reserve\\
Receipt bytes and root growth & repeated local exits & Shows whether receipts
become datum bloat\\
Watcher response time & stale close then contest & Tests the liveness bridge\\
\bottomrule
\end{longtable}

\subsection{Aiken Validator Sketch}

Prototype an Aiken validator for the Head-local factory UTxO. It should check
domain separation, state-number monotonicity, touched-leaf root updates, release
receipts, and a simple ADA-only conservation rule.

\subsection{Virtual-channel Update Traces}

Generate traces for cooperative update, stale local exit attempt, newer-state
materialisation, rebalance between two virtual channels, release receipt
creation, and failed non-interference proof.

\subsection{Close and Fanout Test Cases}

Test Head closure with all virtual channels materialised, a stale Head snapshot
missing materialisation followed by contest with a newer snapshot, excessive
materialisation output count, and a large inline datum. Native-token output
cases belong to the multi-asset extension suite, where rejection under V0-ADA
and later multi-asset fanout behaviour can be tested separately.

\subsection{Non-interference Tests}

Positive tests should show that a local exit touching Alice and Bob does not
affect Carol's balance, reserve claim, membership, or exit path. Negative tests
should show that a balance-only proof is rejected when shared reserve changes.

\subsection{Extension Asset Conservation Matrix}

The V0 acceptance path tests lovelace and min-ADA only. The extension suite
should later test one native asset, multiple native assets under the same
policy, multiple policy IDs, zero-quantity edge cases, and explicit mint/burn
boundaries.

\section{Conclusion}

Lerna Protocol is best understood as a Hydra-native factory discipline. It does
not try to replace Hydra's Head lifecycle, does not compete with Hydrozoa, and
does not rebrand eltoo for Cardano. Its purpose is narrower and more useful:
compactly manage many virtual channels inside Hydra-family Heads while
preserving deterministic materialisation, latest-state settlement, asset
conservation, and factory-local non-interference.

The most promising first implementation is V0. Keep Lerna entirely inside an
open Hydra Head. Use a compact factory UTxO. Require unresolved virtual rights
to materialise before Head close and fanout. Add release receipts so compact
rights cannot be double-spent after materialisation. Treat V1 L1 claim UTxOs as
future research until validators and proof sizes are credible.

If this direction is correct, Lerna is a candidate underexplored layer in the
Cardano scaling stack: Hydra gives fast shared eUTxO execution; Hydrozoa and
Interhead work broaden the Head-family design space; Lerna supplies a
factory-local semantics for making virtual channels safer, compact, and
auditable inside that ecosystem.

\appendix
\section{Hydra Transaction-flow Trace}

This appendix gives the operational V0 trace that a Hydra implementer should
expect to reproduce in a local demo. It is intentionally phrased as a sequence
of Head-visible actions rather than as an abstract state-machine diagram.

\paragraph{Minimal V0 prototype target.}
The smallest useful demo is a local Hydra Head with one compact factory UTxO,
two ADA-only virtual channels, one local exit, one release receipt, deterministic
materialisation before fanout, and a stale snapshot contest scenario. Native
assets, dynamic membership, and V1 claim UTxOs should remain outside the first
demo. The two-channel case should be chosen so that one right materialises while
at least one other right remains compact, since that is the smallest setting
that exercises both release receipts and non-interference boundaries.

\begin{enumerate}[label=\textbf{H\arabic*.},leftmargin=2.6em]
\item \textbf{Open or join a Hydra Head.} The Head participant set, Head id,
contestation period, and node topology are established by the Hydra workflow.
Lerna records the Head id in its factory configuration but does not alter the
Head opening protocol.

\item \textbf{Deposit funds into the Head.} Participants or operators deposit
the UTxOs that will fund the Lerna factory reserve. The deposited outputs must
already respect Cardano value, datum, reference-script, and min-ADA
constraints.

\item \textbf{Create the factory UTxO inside the Head.} A Head-local
transaction creates the compact Lerna factory UTxO, or an equivalent
validator-compatible state object for the prototype environment. Its datum binds
\(\mathsf{factoryId}\), \(\mathsf{factoryEpoch}\), the Head id, participant
registry root, asset registry root, reserve policy, and settlement descriptor.

\item \textbf{Open virtual channels.} Head-local transactions, or off-chain
states later checkpointed into the Head, create virtual-channel rights inside
the factory. Each opening updates the balance, reserve, membership, and
subchannel commitments under the access-manifest rules.

\item \textbf{Update virtual states.} Virtual participants exchange
domain-separated latest-state headers. A deployment may checkpoint every update
into the Head, or checkpoint only when local exit, rebalance, or liveness
pressure requires it.

\item \textbf{Request local exit.} A participant presents the latest virtual
state, access manifest, touched leaves, required signatures, and either a
non-interference proof or full factory authorisation.

\item \textbf{Materialise Head UTxOs.} The Head-local factory validator, or
validator-compatible transaction rule, accepts a transaction that consumes the
compact factory UTxO, creates ordinary Head UTxOs for the exited rights, writes
release receipts into the new compact factory state, and preserves ADA
conservation in the first profile.

\item \textbf{Prepare for shutdown.} If the Head is expected to close, the
factory stops accepting new compact rights once the materialisation threshold is
crossed. Remaining rights are unfolded by deterministic prefix materialisation
until no unresolved unilateral claim remains in the compact state.

\item \textbf{Close the Head.} A Head participant closes with a Hydra snapshot.
If the snapshot is stale with respect to Lerna, an honest watcher must contest
with a newer Hydra snapshot containing the latest Lerna checkpoint or
materialisation.

\item \textbf{Fan out.} Hydra fanout distributes ordinary Head UTxOs. In V0,
the fanned-out UTxO set should contain materialised Lerna outputs or a
cooperatively closed factory state, not unresolved unilateral virtual claims.
\end{enumerate}

\section{Prototype Validator Checklist}

A minimal V0 prototype should report, for each Head-local factory transaction,
which checks were enforced by validator or validator-compatible transaction
rules, which were enforced by off-chain construction, and which remained
test-only assertions.

\paragraph{Validator or transaction-rule enforced.}
\begin{itemize}[leftmargin=2.2em]
\item factory id and Head id binding;
\item factory state-number monotonicity;
\item domain-separated signatures over the latest-state header;
\item access-manifest hash and transition-family binding;
\item touched-root recomputation;
\item release receipt creation and consumption;
\item ADA-only conservation in the first prototype;
\item rejection of non-ADA values under \(\mathsf{V0\mbox{-}ADA}\).
\end{itemize}

\paragraph{Off-chain construction enforced.}
\begin{itemize}[leftmargin=2.2em]
\item selection of the canonical materialisation prefix;
\item min-ADA sufficiency of all materialised outputs;
\item transaction-size, datum-size, witness-size, and execution-budget preflight;
\item watcher/operator routing to an authorised Head participant;
\item operator-budget separation from participant balances.
\end{itemize}

\paragraph{Test-only until benchmarked.}
\begin{itemize}[leftmargin=2.2em]
\item reduced-signature non-interference proof; the first prototype may use
full-authorisation fallback instead;
\item native-asset conservation by policy ID and asset name for extension
profiles;
\item large datum rejection;
\item stale snapshot contest with a newer Lerna checkpoint;
\item measured time-to-materialise under the chosen contestation period;
\item release-receipt replay attempts across factory epochs and batch indices.
\end{itemize}

\section{Open Questions}

\begin{enumerate}[label=\textbf{Q\arabic*.},leftmargin=2.5em]
\item How should virtual participants who are not Head participants delegate
watching and contest authority without turning operators into custodians?
\item What realistic materialisation bound can be achieved under ordinary
Hydra contestation windows for ADA-only outputs, then for native-asset bundles?
\item Which commitment scheme offers the best proof-size and validator-cost
trade-off for touched leaves and non-interference proofs?
\item Can a V1 L1 claim UTxO enforce state monotonicity, signatures,
non-interference, min-ADA, and native-asset conservation within practical
Aiken/Plutus budgets?
\item Would Hydrozoa-style dynamic membership change the assumptions behind the
factory participant registry and rights-dependency graph?
\item How should release receipts be encoded so that they prevent double claims
without causing datum bloat in long-lived factories?
\item What is the minimal V0 factory datum that remains auditable without
making Head-local transactions unwieldy?
\end{enumerate}

\begin{thebibliography}{99}
\raggedright

\bibitem{hydra-paper}
S. Chakravarty, M. D. Coretti, M. Fitzi, P. Gazi, P. Kant, A. Kiayias, and A.
Russell.
\newblock Hydra: Fast Isomorphic State Channels.
\newblock IACR Cryptology ePrint Archive, Report 2020/299.
\newblock \url{https://eprint.iacr.org/2020/299.pdf}.

\bibitem{hydra-docs}
Cardano Scaling.
\newblock Hydra Head protocol documentation.
\newblock \url{https://hydra.family/head-protocol/docs/protocol-overview}.

\bibitem{hydra-repo}
Cardano Scaling.
\newblock Hydra implementation repository.
\newblock \url{https://github.com/cardano-scaling/hydra}.

\bibitem{hydra-known}
Cardano Scaling.
\newblock Hydra known issues and limitations.
\newblock Version 2.0.0 documentation, consulted 5 June 2026.
\newblock \url{https://hydra.family/head-protocol/docs/known-issues}.

\bibitem{hydra-pay}
Obsidian Systems.
\newblock Hydra Pay repository.
\newblock \url{https://github.com/cardano-scaling/hydra-pay}.

\bibitem{hydrozoa}
Cardano Hydrozoa.
\newblock Hydrozoa repository and protocol materials; local repository marks
the project as pre-alpha and under active development.
\newblock \url{https://github.com/cardano-hydrozoa/hydrozoa}.

\bibitem{hydrozoa-spec}
G. Flerovsky, I. Rodionov, and P. Dragos.
\newblock Hydrozoa: Lightweight state channels for Cardano.
\newblock Technical specification working draft, 7 November 2025.
\newblock \url{https://cardano-hydrozoa.github.io/hydrozoa/hydrozoa.pdf}.

\bibitem{interhead}
M. Jourenko, M. Larangeira, and K. Tanaka.
\newblock Interhead Hydra: Two Heads are Better than One.
\newblock IACR Cryptology ePrint Archive, Report 2021/1188.
\newblock \url{https://eprint.iacr.org/2021/1188.pdf}.

\bibitem{aiken}
Aiken.
\newblock A modern smart contract platform for Cardano.
\newblock \url{https://github.com/aiken-lang/aiken}.

\bibitem{cardano-eutxo}
Cardano Documentation.
\newblock The Extended UTxO accounting model.
\newblock \url{https://docs.cardano.org/about-cardano/learn/eutxo-explainer/}.

\bibitem{eltoo}
C. Decker, R. Russell, and O. Osuntokun.
\newblock eltoo: A Simple Layer2 Protocol for Bitcoin.
\newblock \url{https://blockstream.com/eltoo.pdf}.

\bibitem{factories}
C. Burchert, C. Decker, and R. Wattenhofer.
\newblock Scalable Funding of Bitcoin Micropayment Channel Networks.
\newblock \url{https://tik-old.ee.ethz.ch/file//49218d6b9325ddb4f18aef8830ec567b/1/scalable_funding.pdf}.

\bibitem{perun}
S. Dziembowski, S. Faust, and K. Hostakova.
\newblock General State Channel Networks.
\newblock ACM CCS 2018.

\end{thebibliography}

\end{document}
